---
title: "L1-CRISPRi organoids: TE pseudobulk sizefactors"
output: html_notebook
---

This markdown relates to the visualization of pseudobulk quantification of TEs in day 15 cerebral organoids.

We need to calculate sizeFactors per pseudobulked clusters to normalize TEs using unique mapping (crispri_org_pseudobulk_uniq.Rmd).

## Functions

Handy functions to plot featureplots in pseudobulks

```{r}
library(data.table)
library(openxlsx)
library(DESeq2)
library(tidyverse)
library(Seurat)
library(RColorBrewer)
library(stringr)
library(ggpubr)
library(ggplot2)
library(ggpubr)

blues <- brewer.pal(9, 'Blues')[3:9]
pseudobulk_gene <- function(df, gene, minimum = NA, maximum = NA, midpoint = NA, scaled = T){
  tmp <- df[which(df$gene_name == gene),-1]
  rownames(tmp) <- tmp[,ncol(tmp)]
  tmp <- tmp[,-ncol(tmp)]
  tmp <- t(tmp)
  # umap_tmp <- umap[which(umap$region == region),]
  tmp <- merge(crispri_umap, tmp, by.x="sample_cluster", by.y="row.names")
  if(scaled){
    minimum = ifelse(is.na(minimum), log2(min(tmp[,gene])+1), minimum)
    maximum = ifelse(is.na(maximum), log2(max(tmp[,gene])+1) - 2*(sd(log2((tmp[,gene])+1))), maximum)
    midpoint = maximum/2
    tmp[,gene] <- ifelse(tmp[,gene] > maximum, maximum, tmp[,gene])
    plt <- tmp %>%
      arrange(get(gene)) %>%
      ggplot(aes(x=umap_1, y=umap_2, color=log2(get(gene)+1))) + geom_point(size=0.4) 
  }else{
    minimum = ifelse(is.na(minimum), min(tmp[,gene]), minimum)
    maximum = ifelse(is.na(maximum), max(tmp[,gene]) - 2*(sd((tmp[,gene]))), maximum)
    midpoint = maximum/2
    tmp[,gene] <- ifelse(tmp[,gene] > maximum, maximum, tmp[,gene])
    plt <- tmp %>%
      arrange(get(gene)) %>%
      ggplot(aes(x=umap_1, y=umap_2, color=get(gene))) + geom_point(size=0.4) 
  }
  plt <- plt + theme_bw() + 
    scale_colour_gradient2(limits = c(minimum, maximum), high = "#1e386c", low = "#bfdbf0", mid = "#0394cb", midpoint = midpoint) +
    labs(color = gene) +
    facet_wrap(.~condition) +
    ggtitle(gene) +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold"))
  return(plt)
}

pseudobulk_TE <- function(df, TE, minimum = NA, maximum = NA, midpoint = NA, scaled=T){
  tmp <- df[which(df$TE_subfamily == TE),]
  rownames(tmp) <- tmp[,ncol(tmp)]
  tmp <- tmp[,-ncol(tmp)]
  tmp <- t(tmp)
  tmp <- merge(crispri_umap, tmp, by.x="sample_cluster", by.y="row.names")
  if(scaled){
    minimum = ifelse(is.na(minimum), min(log2(tmp[,TE]+1)), minimum)
    maximum = max(log2(tmp[,TE]+1))#ifelse(is.na(maximum), log2(max(tmp[,TE])+1) - 2*(sd(log2((tmp[,TE])+1))), maximum)
    midpoint = minimum + ((maximum - minimum)/2)
    tmp[,TE] <- ifelse(log2(tmp[,TE]+1) > maximum, maximum, log2(tmp[,TE]+1))
    plt <- tmp %>%
      arrange(get(TE)) %>%
      ggplot(aes(x=umap_1, y=umap_2, color=get(TE))) + geom_point(size=0.4) 
  }else{
    minimum = min(tmp[,TE])#ifelse(is.na(minimum), min(tmp[,TE]), minimum)
    maximum =  max(tmp[,TE])#ifelse(is.na(maximum), max(tmp[,TE]) - 2*(sd(tmp[,TE])), maximum)
    midpoint = maximum/2
    tmp[,TE] <- ifelse(tmp[,TE] > maximum, maximum, tmp[,TE])
    plt <- tmp %>%
      arrange(get(TE)) %>%
      ggplot(aes(x=umap_1, y=umap_2, color=get(TE))) + geom_point(size=0.4) 
  }
  plt <- plt + theme_bw() + 
    facet_wrap(.~condition) +
    scale_colour_gradient2(limits = c(minimum, maximum), high = "#1e386c", low = "#bfdbf0", mid = "#0394cb", midpoint = midpoint) +
    labs(color = TE) +
    ggtitle(TE) +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold"))
  return(plt)
}


```

## Read data

TEtranscripts' gene quantification per pseudobulked clusters.

```{r}
path <- "/Volumes/MyPassport/CRISPRi_L1s/10X/3_combinedUMAP_perCluster_res0.1/clusterPipeline/TEcounts/multiple/"
counts <- list()

files <- list.files(path, recursive = T)
for(i in 1:length(files)){
  file <- files[i]
  if(i == 1){
    counts <- fread(paste(path, file, sep="/"), data.table = F)  
    colnames(counts)[2] <- str_remove_all(file, pattern = "_.cntTable")
  }else{
    tmp <- fread(paste(path, file, sep="/"), data.table = F)  
    colnames(tmp)[2] <- str_remove_all(file, pattern = "_.cntTable")
    if(all(counts$`gene/TE` == tmp$`gene/TE`)){
      counts <- cbind(counts, tmp[,2,drop=F])
    }else{
      counts <- merge(counts, tmp, by="gene/TE")
    }
  }
}
rownames(counts) <- counts$`gene/TE`

```

## Subset count matrices to just gene counts and set samplesheet

```{r}
gene_counts <- list()
gene_counts <- counts[which(startsWith(counts$`gene/TE`, "ENSG")),]

samplesheet <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/samplesheet_L1crispr_10X.tab", data.table = F)
samplesheet$sample_name <- paste("crispri_", samplesheet$sample, sep="")

sample_cluster <- data.frame(cluster = colnames(gene_counts)[-1])

sample_cluster$cluster_number <- sapply(str_split(sample_cluster$cluster, "_"), tail, 1)
sample_cluster$sample_name <- gsub('.{2}$', '', sample_cluster$cluster)
# sample_cluster$sample_name <- ifelse(endsWith(sample_cluster$sample_name, "_"), gsub('.{1}$', '', sample_cluster$sample_name), sample_cluster$sample_name)

samplesheet <- merge(samplesheet, sample_cluster, by="sample_name")
rownames(samplesheet) <- samplesheet$cluster

unique(samplesheet$sample_name)
samplesheet_g3 <- samplesheet[which(samplesheet$guide %in% c("gRNA1", "LacZ")),]

```

## Calculate sizefactors per cluster and write to file

```{r}
gene_annotation <- fread("/Volumes/MyPassport/annotations/human/gencode/v38/gencode.v38.annotation.gene_names.tab", skip=1, header = F)
colnames(gene_annotation) <- c("gene_id", "gene_name", "gene_type")
gene_dds <- list()
gene_res <- list()
gene_res_df <- list()

for(i in as.character(unique(samplesheet_g3$cluster_number))){
  samplesheet_tmp <- samplesheet_g3[which(samplesheet_g3$cluster_number == i),]
  gene_dds[[i]] <- DESeqDataSetFromMatrix(gene_counts[,rownames(samplesheet_tmp)], samplesheet_tmp, design =  ~ guide)
  gene_dds[[i]]$guide <- relevel(gene_dds[[i]]$guide, "LacZ")
  gene_dds[[i]] <- DESeq(gene_dds[[i]])
  gene_res[[i]] <- lfcShrink(gene_dds[[i]], "guide_gRNA1_vs_LacZ")
  gene_res_df[[i]] <- as.data.frame(gene_res[[i]])
  gene_res_df[[i]] <- merge(gene_res_df[[i]], gene_annotation[,1:2], by.x="row.names", by.y="gene_id")
  print(i)
}

sizeFactors <- data.frame()
gene_res_df_bound <- data.frame()
for(i in (names(gene_dds))){
  sizeFactors <- rbind(sizeFactors, data.frame(sizeFactor = gene_dds[[i]]$sizeFactor,
                                               cluster = i,
                                               samples = names(gene_dds[[i]]$sizeFactor)))
  gene_res_df[[i]]$cluster <- i
  gene_res_df_bound <- rbind(gene_res_df_bound, gene_res_df[[i]])
} 
openxlsx::write.xlsx(sizeFactors, "/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/gene_sizeFactors_pseudobulk.xlsx")
```


Give some examples from upregulated genes in cluster 0 and 1
```{r}
gene_counts_norm <- gene_counts[,rownames(samplesheet_g3)]
sizeFactors <- sizeFactors[rownames(samplesheet_g3),]
gene_counts_norm[] <- mapply('/', gene_counts_norm[,rownames(samplesheet_g3)], sizeFactors[rownames(samplesheet_g3),"sizeFactor"])
gene_counts_norm <- merge(gene_counts_norm, gene_annotation, by.x="row.names", by.y="gene_id")

pseudobulk_gene_box <- function(gene){
  gene_examples_g3 <- reshape2::melt(gene_counts_norm[which(gene_counts_norm$gene_name %in% gene),c("gene_name", rownames(samplesheet_g3))])
  gene_examples_g3 <- merge(gene_examples_g3, samplesheet_g3, by.x="variable", by.y="cluster")
  gene_examples_g3$guide <- factor(gene_examples_g3$guide, levels = c("LacZ", "gRNA1"))
  tmp <- gene_res_df_bound[which(gene_res_df_bound$gene_name %in% c(gene)),]
  tmp$group1 <- "gRNA1"
  tmp$group2 <- "LacZ"
  tmp$`.y.` <- "value"
  tmp$p.format <- format(tmp$pvalue, digits = 3, scientific = T)
  tmp$log2FoldChange <- round(tmp$log2FoldChange, digits = 3)
  colnames(tmp)[which(colnames(tmp) == "cluster")] <- "cluster_number"
  
  return(gene_examples_g3 %>% 
    filter(gene_name == gene) %>% 
    ggplot(aes(x=guide, y = value)) + 
    geom_jitter(aes(color = guide), width = 0.1, size=0.4) + 
    geom_boxplot(aes(fill = guide), outliers = F, alpha=0.5) + facet_wrap(.~cluster_number, ncol=8) + 
    stat_pvalue_manual(tmp, y.position = 0.95*max(gene_examples_g3[which(gene_examples_g3$gene_name == gene), "value"]), label = "p.format") + 
    theme_bw() + labs(x="", y="Normalized pseudobulk expression") + ggtitle(gene))
}


pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/boxplot_pseudobulk_examples_fig6.pdf", width = 10, height = 3)
pseudobulk_gene_box(c("TOX"))
pseudobulk_gene_box(c("AKAP6"))
pseudobulk_gene_box(c("ADCY2"))
dev.off()
```


