---
title: "L1-CRISPRi organoids: Cell cycle and commonly dysregulated genes"
output: html_notebook
---

This markdown relates to the visualization of day 15 cerebral organoids.

Main questions:
1. Cell cycle
2. Biomarkers 
3. Characterize the present clusters
4. Commonly down / up regulated genes overrepresentation test

## Read data and add metadata

Including previously defined cell types (see crispri_org.Rmd), batch, cell line, condition, etc.

```{r}
library(data.table)
library(openxlsx)
library(tidyverse)
library(Seurat)
library(stringr)
library(ggpubr)
library(ggplot2)
library(ggpubr)

crispri_org <- readRDS("/Volumes/MyPassport/CRISPRi_L1s/10X/crispri.rds")

samples_metadata <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/samplesheet_L1crispr_10X.tab", data.table = F)
metadata <- FetchData(crispri_org, "orig.ident")
metadata$barcode <- rownames(metadata)
metadata <- merge(metadata, samples_metadata, by.y="sample", by.x="orig.ident")
rownames(metadata) <- metadata$barcode
crispri_org <- AddMetaData(crispri_org, metadata[,c("guide", "batch", "cell_line", "seqnum")], col.name = c("condition", "batch", "cell_line", "seqnum"))
coldata_celltypes <- reshape2::melt(list("NPCs" = as.character(c(0:2, 7)),
                    "Early neurons" = as.character(c(4,5)),
                    "Endothelial" = as.character(c(3,6))))

crispri_org_join <- JoinLayers(crispri_org)

### ADD cell type to metadata
clusters_per_cell <- FetchData(crispri_org_join, "seurat_clusters")
clusters_per_cell$barcode <- rownames(clusters_per_cell)
celltypes_per_cell <- merge(coldata_celltypes, clusters_per_cell, by.x="value", by.y="seurat_clusters")
colnames(celltypes_per_cell)[2] <- "celltype"
rownames(celltypes_per_cell) <- celltypes_per_cell$barcode
crispri_org_join <- AddMetaData(crispri_org_join, metadata = celltypes_per_cell[,c("celltype"), drop=F], col.name = "celltype")

cellline_per_cell <- FetchData(crispri_org_join, c("condition", "cell_line", "batch", "celltype", "seurat_clusters"))
cellline_per_cell$barcode <- rownames(cellline_per_cell)
cellline_per_cell$condition_cell_line <- paste(cellline_per_cell$condition, cellline_per_cell$cell_line, sep="_")
cellline_per_cell$condition_cluster <- paste(cellline_per_cell$condition, cellline_per_cell$seurat_clusters, sep="_")
cellline_per_cell$condition_celltype <- paste(cellline_per_cell$condition, cellline_per_cell$celltype, sep="_")
cellline_per_cell$condition_cell_line_batch <- paste(cellline_per_cell$condition, cellline_per_cell$cell_line, cellline_per_cell$batch, sep="_")
crispri_org_join <- AddMetaData(crispri_org_join, metadata = cellline_per_cell[,c("condition_cell_line", "condition_cell_line_batch", "condition_celltype"), drop=F], col.name = c("condition_cell_line", "condition_cell_line_batch", "condition_celltype"))
cellline_per_cell$condition_cell_line_cluster <- paste(cellline_per_cell$condition_cell_line, cellline_per_cell$seurat_clusters, sep = "_")
cellline_per_cell$condition_cell_line_batch_cluster <- paste(cellline_per_cell$condition_cell_line_batch, cellline_per_cell$seurat_clusters, sep = "_")
cellline_per_cell$condition_cell_line_celltype <- paste(cellline_per_cell$condition_cell_line, cellline_per_cell$celltype, sep = "_")
cellline_per_cell$condition_cell_line_batch_celltype <- paste(cellline_per_cell$condition_cell_line_batch, cellline_per_cell$celltype, sep = "_")
crispri_org_join <- AddMetaData(crispri_org_join, metadata = cellline_per_cell[,c("condition_cell_line_batch_cluster", "condition_cell_line_batch_celltype", "condition_cluster", "condition_cell_line_cluster", "condition_cell_line_celltype"), drop=F], col.name = c("condition_cell_line_batch_cluster", "condition_cell_line_batch_celltype", "condition_cluster", "condition_cell_line_cluster", "condition_cell_line_celltype"))
```

## Cell cycle

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
crispri_org_join <- CellCycleScoring(crispri_org_join, s.features = s.genes, g2m.features = g2m.genes, set.ident = F, assay="RNA")
crispri_org_join <- AddMetaData(crispri_org_join, metadata = ifelse(crispri_org_join$Phase == "G1", "non-cycling", "cycling"), col.name = "cellCycle")

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/dotplot_crispri_org_markers_clusters.pdf")
DotPlot(crispri_org_join, features = unique(c("G2M.Score","S.Score",  "TJP1",  "GLI3", "CUX2", "PAX6", "SOX2", "PCNA", "EOMES", "BCL11B", "MAP2", "NCAM1", "RBFOX3", "TBR1", "DCX", "GAD1")), group.by = "seurat_clusters", assay = "RNA") & theme(axis.text.x = element_text(angle=90, hjust=1)) & coord_flip()
# dev.off()

cellcycle_scores <- FetchData(tmp, c("umap_1", "umap_2","day", "G2M.Score", "S.Score", "condition", "seurat_clusters"))
cellcycle_scores <- merge(cellcycle_scores, FetchData(tmp, c("celltype", "condition_cell_line_batch")), by="row.names")
rownames(cellcycle_scores) <- cellcycle_scores$Row.names
cellcycle_scores <- cellcycle_scores[,-1]

cellcycle_scores$score <- cellcycle_scores$G2M.Score + cellcycle_scores$S.Score
cellcycle_scores <- cellcycle_scores[order(cellcycle_scores$seurat_clusters, cellcycle_scores$condition),]

# Featureplot coloured by cell cycle score
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_crispri_org_cellcycle_score_hiPS6.pdf", width = 5, height = 4)
ggplot(cellcycle_scores, aes(x=umap_1, y=umap_2, colour=score)) + geom_point(size=0.5)+
  scale_colour_gradient2(low="white", mid="lightgrey", high = "red") + theme_bw() +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold"))
# dev.off()


## Cell cycle visualization per experiment
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_crispri_org_cellcycle_score.pdf", width = 14, height = 8)
ggarrange(
  ggplot(cellcycle_scores, aes(x=umap_1, y=umap_2, colour=G2M.Score)) + geom_point(size=0.5) + facet_wrap(.~condition_cell_line_batch, ncol = 8) + scale_colour_gradient2(low="white", mid="lightgrey", high = "red") + theme_bw() +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold")),
          
  ggplot(cellcycle_scores, aes(x=umap_1, y=umap_2, colour=S.Score)) + geom_point(size=0.5) + facet_wrap(.~condition_cell_line_batch, ncol = 8) + scale_colour_gradient2(low="white", mid="lightgrey", high = "red") + theme_bw() +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold")), ncol=1)
# dev.off()
```

## Markers per experiment per cell type

```{r}
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/dotplot_crispri_org_markers_celltypes_lacz.pdf", height = 6, width = 7)
DotPlot(subset(crispri_org_join, condition == "LacZ" & cell_line == "hiPS6"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org (all)")
DotPlot(subset(crispri_org_join, condition == "LacZ" & cell_line == "hiPS48"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS48 org (all)")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS6_1"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org batch 1")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS6_2"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org batch 2")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS6_3"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org batch 3")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS6_4"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org batch 4")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS48_1"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS48 org batch 1")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS48_2"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS48 org batch 2")
# dev.off()

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/dotplot_crispri_org_markers_celltypes.pdf", height = 5, width = 7)
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_1", "gRNA1_hiPS6_1")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g1 batch 1")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_2", "gRNA1_hiPS6_2")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g1 batch 2")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_2", "gRNA2_hiPS6_2")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g2 batch 2")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_3", "gRNA1_hiPS6_3")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g1 batch 3")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_4", "gRNA1_hiPS6_4")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g1 batch 4")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_4", "gRNA2_hiPS6_4")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g2 batch 4")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS48_1", "gRNA1_hiPS48_1")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS48 org g1 batch 1")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS48_2", "gRNA1_hiPS48_2")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS48 org g1 batch 2")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS48_2", "gRNA2_hiPS48_2")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS48 org g2 batch 2")
# dev.off()
```

## Characterize the clusters (DEA per cluster)

Top 10 differentially expressed genes per cluster
```{r}
set.seed(10)
crispri_org_join_markers <- FindAllMarkers(crispri_org_join, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

gene_annotation <- fread("/Volumes/MyPassport/annotations/human/gencode/v38/gencode.v38.annotation.gene_names.tab", data.table=F, skip=1, header =F )[,-1]
colnames(gene_annotation) <- c("gene_name", "gene_type")
crispri_org_join_markers <- merge(crispri_org_join_markers, gene_annotation, by.x="gene", by.y="gene_name")
crispri_org_join_markers_top10 <- crispri_org_join_markers %>% 
  filter(gene_type == "protein_coding") %>% 
  filter(pct.1 > 0.4) %>% 
  group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC) %>% 
  arrange(cluster)

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/crispri_org_markers_dotplot.pdf", height = 12, width = 7)
DotPlot(crispri_org, features = crispri_org_join_markers_top10$gene, group.by  = "seurat_clusters", assay = "RNA", cols = "RdBu") & coord_flip() 
# dev.off()

# write.xlsx(crispri_org_join_markers, "/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/crispri_org_markers_clusters.xlsx")

```

## Log2FC of consistently upregulated or downregulated genes

### Upregulated

So, the files read here are based on the results of hiPS6 organoids. We need to find the genes that have at least an increase of 20% (log2(1.2) = 0.26) in the base experiment (g1 hiPS6), and the same direction in the rest of the experiments (log2FC > 0).

```{r}
deas_clusters <- readRDS("/Volumes/MyPassport/CRISPRi_L1s/10X/src/r_scripts/crispri_org_deas_clusters.rds")

upreg_0 <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/upreg_cluster_0.txt", header = F, data.table = F)
upreg_1 <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/upreg_cluster_1.txt", header = F, data.table = F)
downreg_0 <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/downreg_cluster_0.txt", header = F, data.table = F)
downreg_1 <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/downreg_cluster_1.txt", header = F, data.table = F)

# At least an increase of 20% (log2(1.2) = 0.26) in the base (g1 hiPS6), and the same direction in the rest of the experiments
upreg_0_threshold <- upreg_0[which((deas_clusters$g1_hiPS6$`0`[upreg_0$V1, "0_g1_hiPS6_avg_log2FC"] > 0.25) &
                                     deas_clusters$g2_hiPS6$`0`[upreg_0$V1, "avg_log2FC"] > 0 &
                                     deas_clusters$g1_hiPS48$`0`[upreg_0$V1, "avg_log2FC"] > 0 &
                                     deas_clusters$g2_hiPS48$`0`[upreg_0$V1, "avg_log2FC"] > 0),]

upreg_1_threshold <- upreg_1[which((deas_clusters$g1_hiPS6$`1`[upreg_1$V1, "1_g1_hiPS6_avg_log2FC"] > 0.25) &
                                     deas_clusters$g2_hiPS6$`1`[upreg_1$V1, "avg_log2FC"] > 0 &
                                     deas_clusters$g1_hiPS48$`1`[upreg_1$V1, "avg_log2FC"] > 0 &
                                     deas_clusters$g2_hiPS48$`1`[upreg_1$V1, "avg_log2FC"] > 0),]

upreg_both <- unique(upreg_0_threshold[which(upreg_0_threshold %in% upreg_1_threshold)], 
                     upreg_1_threshold[which(upreg_1_threshold %in% upreg_0_threshold)])

```

## Heatmap with upregulated genes 

Average per cell type

```{r}
upreg_both_df <- FetchData(crispri_org, c("orig.ident","condition_cell_line_celltype", upreg_both))

upreg_both_df_mean_sample <- aggregate(upreg_both_df[,upreg_both], by=list(upreg_both_df$orig.ident, upreg_both_df$condition_cell_line_celltype), FUN=mean)
upreg_both_df_mean_sample$group <- paste(upreg_both_df_mean_sample$Group.1, upreg_both_df_mean_sample$Group.2)
rownames(upreg_both_df_mean_sample) <- upreg_both_df_mean_sample$group
upreg_both_df_mean_sample <- t(upreg_both_df_mean_sample[,upreg_both])

library(pheatmap)
colannot_endothelial <- data.frame(sample=sapply(str_split(colnames(upreg_both_df_mean_sample_endothelial), " "), `[[`, 1),
           colname = colnames(upreg_both_df_mean_sample_endothelial))
colannot_endothelial <- merge(colannot_endothelial, samples_metadata )
rownames(colannot_endothelial) <- colannot_endothelial$colname

colannot_npcs <- data.frame(sample=sapply(str_split(colnames(upreg_both_df_mean_sample_npcs), " "), `[[`, 1),
           colname = colnames(upreg_both_df_mean_sample_npcs))
colannot_npcs <- merge(colannot_npcs, samples_metadata )
rownames(colannot_npcs) <- colannot_npcs$colname

colannot_neurons <- data.frame(sample=sapply(str_split(colnames(upreg_both_df_mean_sample_neurons), " "), `[[`, 1),
           colname = colnames(upreg_both_df_mean_sample_neurons))
colannot_neurons <- merge(colannot_neurons, samples_metadata )
rownames(colannot_neurons) <- colannot_neurons$colname

upreg_both_df_mean_sample_npcs <- upreg_both_df_mean_sample[,colnames(upreg_both_df_mean_sample)[which(endsWith(colnames(upreg_both_df_mean_sample), "NPCs"))]]

upreg_both_df_mean_sample_neurons <- upreg_both_df_mean_sample[,colnames(upreg_both_df_mean_sample)[which(endsWith(colnames(upreg_both_df_mean_sample), "neurons"))]]

upreg_both_df_mean_sample_endothelial <- upreg_both_df_mean_sample[,colnames(upreg_both_df_mean_sample)[which(endsWith(colnames(upreg_both_df_mean_sample), "Endothelial"))]]

tmp <- pheatmap(log2(upreg_both_df_mean_sample_npcs+1), cluster_cols = F, show_colnames = F, annotation_col = colannot_npcs[colnames(upreg_both_df_mean_sample_npcs),c(3,4)], scale="row", main = "NPCs", color = colorRampPalette(c("blue", "white", "red"))(100), breaks = seq(-4,4,length.out=100))

pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_crispri_org_log2fc025_nopadj_average.pdf", height = 7, width = 5)
pheatmap(log2(upreg_both_df_mean_sample_npcs[tmp$tree_row$labels,]+1), cluster_cols = F, show_colnames = F, annotation_col = colannot_npcs[colnames(upreg_both_df_mean_sample_npcs),c(3,4)], scale="row", main = "NPCs", color = colorRampPalette(c("blue", "white", "red"))(100), breaks = seq(-4,4,length.out=100), cluster_rows = F)

pheatmap(log2(upreg_both_df_mean_sample_neurons[tmp$tree_row$labels,]+1), cluster_cols = F, show_colnames = F, annotation_col = colannot_neurons[colnames(upreg_both_df_mean_sample_neurons),c(3,4)], scale="row", main = "Neurons", color = colorRampPalette(c("blue", "white", "red"))(100), breaks = seq(-4,4,length.out=100), cluster_rows = F)

pheatmap(log2(upreg_both_df_mean_sample_endothelial[tmp$tree_row$labels,]+1), cluster_cols = F, show_colnames = F, annotation_col = colannot_endothelial[colnames(upreg_both_df_mean_sample_endothelial),c(3,4)], scale="row", main = "Endothelial", color = colorRampPalette(c("blue", "white", "red"))(100), breaks = seq(-4,4,length.out=100), cluster_rows = F)

dev.off()
```


### Downregulated

So, the files read here are based on the results of hiPS6 organoids. We need to find the genes that have at least a decrease of 20% (log2(1.2) = 0.26) in the base experiment (g1 hiPS6), and the same direction in the rest of the experiments (log2FC < 0).

```{r}
downreg_0_threshold <- downreg_0[which((deas_clusters$g1_hiPS6$`0`[downreg_0$V1, "0_g1_hiPS6_avg_log2FC"] < -0.25) &
                                     deas_clusters$g2_hiPS6$`0`[downreg_0$V1, "avg_log2FC"] < 0 &
                                     deas_clusters$g1_hiPS48$`0`[downreg_0$V1, "avg_log2FC"] < 0 &
                                     deas_clusters$g2_hiPS48$`0`[downreg_0$V1, "avg_log2FC"] < 0),]

downreg_1_threshold <- downreg_1[which((deas_clusters$g1_hiPS6$`1`[downreg_1$V1, "1_g1_hiPS6_avg_log2FC"] < -0.25) &
                                     deas_clusters$g2_hiPS6$`1`[downreg_1$V1, "avg_log2FC"] < 0 &
                                     deas_clusters$g1_hiPS48$`1`[downreg_1$V1, "avg_log2FC"] < 0 &
                                     deas_clusters$g2_hiPS48$`1`[downreg_1$V1, "avg_log2FC"] < 0),]

downreg_both <- unique(downreg_0_threshold[which(downreg_0_threshold %in% downreg_1_threshold)], 
                     downreg_1_threshold[which(downreg_1_threshold %in% downreg_0_threshold)])
```

## Upregulated genes heatmaps
```{r}
library(pheatmap)
upreg_genes_df <- FetchData(crispri_org_join, c(upreg_both, "condition_cluster","condition", "celltype", "seurat_clusters", "cell_line", "batch"))

# upreg_genes_df <- upreg_genes_df %>% 
#   filter(condition_cluster %in% c("LacZ_0", "LacZ_1", "LacZ_4",
#                                   "gRNA1_0", "gRNA1_1", "gRNA1_4",
#                                   "gRNA2_0", "gRNA2_1", "gRNA2_4")) 

cluster_rows_by_group <- function(df_cluster, group_rows, samplesheet){
  # Calculate eigen values
    df_cluster <- t(df_cluster)
    eigenvalues <- svd(t(scale(df_cluster)),nu=1,nv=1)$v
    scaledExpr <- scale(df_cluster)
    averExpr <- rowMeans(scaledExpr, na.rm = TRUE)
    if(cor(averExpr,eigenvalues) < 0){
      eigenvalues <- -eigenvalues
    }
    index_eigen <- order(eigenvalues)
    index_eigen <- rownames(df_cluster[index_eigen,])
    # Add group variable in samplesheet
    # If more than one, we collapse those columns
    if(length(group_rows) > 1){
      samplesheet$Group <- apply(samplesheet[,group_rows], 1, paste, collapse=' ')    
    }else{
      samplesheet$Group <- samplesheet[,group_rows]
    }
    index_reorder <- c()
    # For each group
    for(eachgroup in unique(samplesheet$Group)){
      index_tempEigen <- index_eigen[index_eigen %in% rownames(samplesheet[which(samplesheet$Group == eachgroup),])]
      # we calculate the sample distance to each cluster of samples (in this case nuclei)
      sampleDist <- dist(df_cluster[index_tempEigen,], method="euclidean")
      sampleClust <- hclust(sampleDist)
      index_clust <- index_tempEigen[sampleClust$order]
      # if(cor(index_clust,c(1:length(index_tempEigen))) < 0){
      #   index_clust <- rev(index_clust)
      # }
      # Final order
      index_reorder <- c(index_reorder,index_clust)
    }
    return(index_reorder)
}


df <- t(upreg_genes_df[,upreg_both])
df <- df[which(rowSums(df) > 0),]
df <- df[,which(colSums(df) > 0)]
index_reorder <- cluster_rows_by_group(df_cluster = df, group_rows = c("seurat_clusters", "celltype", "condition"), samplesheet = upreg_genes_df[,c("seurat_clusters", "celltype", "condition"), drop=F])

upreg_genes_df_col_annot <- upreg_genes_df[colnames(df[upreg_both,index_reorder]),c("condition", "seurat_clusters", "celltype"),drop=F]
upreg_genes_df_col_annot$condition_cluster <- paste(paste(upreg_genes_df_col_annot$celltype, upreg_genes_df_col_annot$seurat_clusters, sep="_"), upreg_genes_df_col_annot$condition, sep="_")
upreg_genes_df_col_annot_split <- split(upreg_genes_df_col_annot, f = c(upreg_genes_df_col_annot$condition_cluster))

index_reorder <- rownames(rbind(upreg_genes_df_col_annot_split$`NPCs_0_LacZ`,
      upreg_genes_df_col_annot_split$`NPCs_0_gRNA1`,
      upreg_genes_df_col_annot_split$`NPCs_0_gRNA2`,
      upreg_genes_df_col_annot_split$`NPCs_1_LacZ`,
      upreg_genes_df_col_annot_split$`NPCs_1_gRNA1`,
      upreg_genes_df_col_annot_split$`NPCs_1_gRNA2`,
      upreg_genes_df_col_annot_split$`NPCs_2_LacZ`,
      upreg_genes_df_col_annot_split$`NPCs_2_gRNA1`,
      upreg_genes_df_col_annot_split$`NPCs_2_gRNA2`,
      upreg_genes_df_col_annot_split$`NPCs_7_LacZ`,
      upreg_genes_df_col_annot_split$`NPCs_7_gRNA1`,
      upreg_genes_df_col_annot_split$`NPCs_7_gRNA2`,
      upreg_genes_df_col_annot_split$`Endothelial_3_LacZ`,
      upreg_genes_df_col_annot_split$`Endothelial_6_LacZ`,
      upreg_genes_df_col_annot_split$`Endothelial_3_gRNA1`,
      upreg_genes_df_col_annot_split$`Endothelial_6_gRNA1`,
      upreg_genes_df_col_annot_split$`Endothelial_3_gRNA2`,
      upreg_genes_df_col_annot_split$`Endothelial_6_gRNA2`,
      upreg_genes_df_col_annot_split$`Early neurons_4_LacZ`,
      upreg_genes_df_col_annot_split$`Early neurons_5_LacZ`,
      upreg_genes_df_col_annot_split$`Early neurons_4_gRNA1`,
      upreg_genes_df_col_annot_split$`Early neurons_5_gRNA1`,
      upreg_genes_df_col_annot_split$`Early neurons_4_gRNA2`,
      upreg_genes_df_col_annot_split$`Early neurons_5_gRNA2`))

gaps <- rep((which(!duplicated(upreg_genes_df[colnames(df[upreg_both,index_reorder]),c("condition", "seurat_clusters"),drop=F]))[-1]-1), 3)

pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_crispri_org_log2fc025.pdf", width = 14, height = 10)
pheatmap(df[upreg_both,index_reorder], annotation_col = upreg_genes_df[colnames(df[upreg_both,index_reorder]),c("condition", "seurat_clusters", "celltype", "cell_line"),drop=F], border_color = NA, show_colnames = F, cluster_cols = F, scale="row", color = colorRampPalette(c("blue", "white", "red"))(100), breaks = seq(-4,4,length.out=100))
dev.off()

png("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_crispri_org_log2fc025.png", units = "in", width = 14, height = 14, res=300)
pheatmap(df[upreg_both,index_reorder], annotation_col = upreg_genes_df[colnames(df[upreg_both,index_reorder]),c("condition", "seurat_clusters", "celltype", "cell_line"),drop=F], border_color = NA, show_colnames = F, cluster_cols = F, scale="row", color = colorRampPalette(c("blue", "white", "red"))(100), breaks = seq(-4,4,length.out=100))
dev.off()
```

```{r}
downreg_genes_df <- FetchData(crispri_org_join, c(downreg_both, "condition_cluster","condition", "celltype", "seurat_clusters", "cell_line", "batch"))

df <- t(downreg_genes_df[,downreg_both])
df <- df[which(rowSums(df) > 0),]
df <- df[,which(colSums(df) > 0)]
index_reorder <- cluster_rows_by_group(df_cluster = df, group_rows = c("seurat_clusters", "celltype", "condition"), samplesheet = downreg_genes_df[,c("seurat_clusters", "celltype", "condition"), drop=F])

downreg_genes_df_col_annot <- downreg_genes_df[colnames(df[downreg_both,index_reorder]),c("condition", "seurat_clusters", "celltype"),drop=F]
downreg_genes_df_col_annot$condition_cluster <- paste(paste(downreg_genes_df_col_annot$celltype, downreg_genes_df_col_annot$seurat_clusters, sep="_"), downreg_genes_df_col_annot$condition, sep="_")
downreg_genes_df_col_annot_split <- split(downreg_genes_df_col_annot, f = c(downreg_genes_df_col_annot$condition_cluster))

index_reorder <- rownames(rbind(downreg_genes_df_col_annot_split$`NPCs_0_LacZ`,
      downreg_genes_df_col_annot_split$`NPCs_0_gRNA1`,
      downreg_genes_df_col_annot_split$`NPCs_0_gRNA2`,
      downreg_genes_df_col_annot_split$`NPCs_1_LacZ`,
      downreg_genes_df_col_annot_split$`NPCs_1_gRNA1`,
      downreg_genes_df_col_annot_split$`NPCs_1_gRNA2`,
      downreg_genes_df_col_annot_split$`NPCs_2_LacZ`,
      downreg_genes_df_col_annot_split$`NPCs_2_gRNA1`,
      downreg_genes_df_col_annot_split$`NPCs_2_gRNA2`,
      downreg_genes_df_col_annot_split$`NPCs_7_LacZ`,
      downreg_genes_df_col_annot_split$`NPCs_7_gRNA1`,
      downreg_genes_df_col_annot_split$`NPCs_7_gRNA2`,
      downreg_genes_df_col_annot_split$`Endothelial_3_LacZ`,
      downreg_genes_df_col_annot_split$`Endothelial_6_LacZ`,
      downreg_genes_df_col_annot_split$`Endothelial_3_gRNA1`,
      downreg_genes_df_col_annot_split$`Endothelial_6_gRNA1`,
      downreg_genes_df_col_annot_split$`Endothelial_3_gRNA2`,
      downreg_genes_df_col_annot_split$`Endothelial_6_gRNA2`,
      downreg_genes_df_col_annot_split$`Early neurons_4_LacZ`,
      downreg_genes_df_col_annot_split$`Early neurons_5_LacZ`,
      downreg_genes_df_col_annot_split$`Early neurons_4_gRNA1`,
      downreg_genes_df_col_annot_split$`Early neurons_5_gRNA1`,
      downreg_genes_df_col_annot_split$`Early neurons_4_gRNA2`,
      downreg_genes_df_col_annot_split$`Early neurons_5_gRNA2`))


png("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/downreg_genes_crispri_org_log2fc025.png", units = "in", width = 14, height = 12, res=300)
pheatmap(df[downreg_both,index_reorder], annotation_col = downreg_genes_df[colnames(df[downreg_both,index_reorder]),c("condition", "seurat_clusters", "celltype", "cell_line"),drop=F], border_color = NA, show_colnames = F, cluster_cols = F, scale="row", color = colorRampPalette(c("blue", "white", "red"))(100), breaks = seq(-4,4,length.out=100))
dev.off()

pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/downreg_genes_crispri_org_log2fc025.pdf", width = 14, height = 12)
pheatmap(df[downreg_both,index_reorder], annotation_col = downreg_genes_df[colnames(df[downreg_both,index_reorder]),c("condition", "seurat_clusters", "celltype", "cell_line"),drop=F], border_color = NA, show_colnames = F, cluster_cols = F, scale="row", color = colorRampPalette(c("blue", "white", "red"))(100), breaks = seq(-4,4,length.out=100))
dev.off()
```

### Only upregulated in a particular cluster

So, from the complete list of upregulated genes in cluster 0, we have filtered:
* OUT: Genes that dont have the same direction of the effect as the base experiment (upreg_0 not in upreg_0_threshold). 
* IN: Genes that are NOT upregulated in cluster 1 in ANY experiment. We want ONLY upreg in cluster 0.
* IN: Genes that are consistently upregulated in cluster 0 in ALL experiments (at least a little bit! log2FC > 0)
```{r}
upreg_0_only <- upreg_0[which(!(upreg_0$V1 %in% upreg_0_threshold) &
                                !(deas_clusters$g1_hiPS6$`1`[upreg_0$V1, "1_g1_hiPS6_avg_log2FC"] > 0) &
                                !(deas_clusters$g2_hiPS6$`1`[upreg_0$V1, "avg_log2FC"] > 0) &
                                !(deas_clusters$g1_hiPS48$`1`[upreg_0$V1, "avg_log2FC"] > 0) &
                                !(deas_clusters$g2_hiPS48$`1`[upreg_0$V1, "avg_log2FC"] > 0) &
                                
                                (deas_clusters$g1_hiPS6$`0`[upreg_0$V1, "0_g1_hiPS6_avg_log2FC"] > 0.25) &
                                (deas_clusters$g2_hiPS6$`0`[upreg_0$V1, "avg_log2FC"] > 0) &
                                (deas_clusters$g1_hiPS48$`0`[upreg_0$V1, "avg_log2FC"] > 0) &
                                (deas_clusters$g2_hiPS48$`0`[upreg_0$V1, "avg_log2FC"] > 0)  ),]

upreg_1_only <- upreg_1[which(!(upreg_1$V1 %in% upreg_1_threshold) &
                                !(deas_clusters$g1_hiPS6$`0`[upreg_1$V1, "0_g1_hiPS6_avg_log2FC"] > 0) &
                                !(deas_clusters$g2_hiPS6$`0`[upreg_1$V1, "avg_log2FC"] > 0) &
                                !(deas_clusters$g1_hiPS48$`0`[upreg_1$V1, "avg_log2FC"] > 0) &
                                !(deas_clusters$g2_hiPS48$`0`[upreg_1$V1, "avg_log2FC"] > 0) &
                                
                                (deas_clusters$g1_hiPS6$`1`[upreg_1$V1, "1_g1_hiPS6_avg_log2FC"] > 0.25) &
                                (deas_clusters$g2_hiPS6$`1`[upreg_1$V1, "avg_log2FC"] > 0) &
                                (deas_clusters$g1_hiPS48$`1`[upreg_1$V1, "avg_log2FC"] > 0) &
                                (deas_clusters$g2_hiPS48$`1`[upreg_1$V1, "avg_log2FC"] > 0)  ),]

downreg_0_only <- downreg_0[which(!(downreg_0$V1 %in% downreg_0_threshold) &
                                !(deas_clusters$g1_hiPS6$`1`[downreg_0$V1, "1_g1_hiPS6_avg_log2FC"] < 0) &
                                !(deas_clusters$g2_hiPS6$`1`[downreg_0$V1, "avg_log2FC"] < 0) &
                                !(deas_clusters$g1_hiPS48$`1`[downreg_0$V1, "avg_log2FC"] < 0) &
                                !(deas_clusters$g2_hiPS48$`1`[downreg_0$V1, "avg_log2FC"] < 0) &
                                
                                (deas_clusters$g1_hiPS6$`0`[downreg_0$V1, "0_g1_hiPS6_avg_log2FC"] < -0.25) &
                                (deas_clusters$g2_hiPS6$`0`[downreg_0$V1, "avg_log2FC"] < 0) &
                                (deas_clusters$g1_hiPS48$`0`[downreg_0$V1, "avg_log2FC"] < 0) &
                                (deas_clusters$g2_hiPS48$`0`[downreg_0$V1, "avg_log2FC"] < 0)  ),]

downreg_1_only <- downreg_1[which(!(downreg_1$V1 %in% downreg_1_threshold) &
                                !(deas_clusters$g1_hiPS6$`0`[downreg_1$V1, "0_g1_hiPS6_avg_log2FC"] < 0) &
                                !(deas_clusters$g2_hiPS6$`0`[downreg_1$V1, "avg_log2FC"] < 0) &
                                !(deas_clusters$g1_hiPS48$`0`[downreg_1$V1, "avg_log2FC"] < 0) &
                                !(deas_clusters$g2_hiPS48$`0`[downreg_1$V1, "avg_log2FC"] < 0) &
                                
                                (deas_clusters$g1_hiPS6$`1`[downreg_1$V1, "1_g1_hiPS6_avg_log2FC"] < -0.25) &
                                (deas_clusters$g2_hiPS6$`1`[downreg_1$V1, "avg_log2FC"] < 0) &
                                (deas_clusters$g1_hiPS48$`1`[downreg_1$V1, "avg_log2FC"] < 0) &
                                (deas_clusters$g2_hiPS48$`1`[downreg_1$V1, "avg_log2FC"] < 0)  ),]


deg <- c(downreg_both, downreg_0_only, downreg_1_only,
         upreg_both, upreg_0_only, upreg_1_only)
```



```{r}
g1_hiPS6_0_deg_df <- deas_clusters$g1_hiPS6$`0`[which(deas_clusters$g1_hiPS6$`0`$gene %in% deg),]
g1_hiPS6_1_deg_df <- deas_clusters$g1_hiPS6$`1`[which(deas_clusters$g1_hiPS6$`1`$gene %in% deg),]
g2_hiPS6_0_deg_df <- deas_clusters$g2_hiPS6$`0`[which(deas_clusters$g2_hiPS6$`0`$gene %in% deg),]
g2_hiPS6_1_deg_df <- deas_clusters$g2_hiPS6$`1`[which(deas_clusters$g2_hiPS6$`1`$gene %in% deg),]
g1_hiPS48_0_deg_df <- deas_clusters$g1_hiPS48$`0`[which(deas_clusters$g1_hiPS48$`0`$gene %in% deg),]
g1_hiPS48_1_deg_df <- deas_clusters$g1_hiPS48$`1`[which(deas_clusters$g1_hiPS48$`1`$gene %in% deg),]
g2_hiPS48_1_deg_df <- deas_clusters$g2_hiPS48$`0`[which(deas_clusters$g2_hiPS48$`0`$gene %in% deg),]
g2_hiPS48_1_deg_df <- deas_clusters$g2_hiPS48$`1`[which(deas_clusters$g2_hiPS48$`1`$gene %in% deg),]

g1_hiPS6_0_deg_df$type <- "g1_hiPS6_0"
g1_hiPS6_1_deg_df$type <- "g1_hiPS6_1"
g2_hiPS6_0_deg_df$type <- "g2_hiPS6_0"
g2_hiPS6_1_deg_df$type <- "g2_hiPS6_1"
g1_hiPS48_0_deg_df$type <- "g1_hiPS48_0"
g1_hiPS48_1_deg_df$type <- "g1_hiPS48_1"
g2_hiPS48_1_deg_df$type <- "g2_hiPS48_1"
g2_hiPS48_1_deg_df$type <- "g2_hiPS48_1"

colnames(g1_hiPS6_0_deg_df)[2] <- "avg_log2FC"
colnames(g1_hiPS6_1_deg_df)[2] <- "avg_log2FC"

cluster_0_1_deg_df <- rbind(g1_hiPS6_0_deg_df,
                            g1_hiPS6_1_deg_df,
                            g2_hiPS6_0_deg_df,
                            g2_hiPS6_1_deg_df,
                            g1_hiPS48_0_deg_df,
                            g1_hiPS48_1_deg_df,
                            g2_hiPS48_1_deg_df,
                            g2_hiPS48_1_deg_df)

unique(cluster_0_1_deg_df$type)
cluster_0_1_deg_df <- unique(cluster_0_1_deg_df)
cluster_0_1_deg_df_dcast <- reshape2::dcast(cluster_0_1_deg_df[,c("avg_log2FC", "type", "gene")], formula = gene~type, value.var = "avg_log2FC")

cluster_0_1_deg_df_dcast[is.na(cluster_0_1_deg_df_dcast)] <- 0
rownames(cluster_0_1_deg_df_dcast) <- cluster_0_1_deg_df_dcast$gene
cluster_0_1_deg_df_dcast <- cluster_0_1_deg_df_dcast[,-1]

bk_neg <- seq(-16,0, length.out=50)
bk_pos <- seq(0,3.5, length.out=50)
bk <- c(bk_neg, bk_pos)
bk <- unique(bk)
color_pos <- colorRampPalette(c("white", "red"))(50)
color_neg <- colorRampPalette(c("blue", "white"))(50)

library(pheatmap)
pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/downreg_genes_cluster_0_1.pdf", height = 15, width = 7)
pheatmap(cluster_0_1_deg_df_dcast[downreg_both, c("g1_hiPS6_0", "g2_hiPS6_0", "g1_hiPS48_0", "g1_hiPS6_1", "g2_hiPS6_1", "g1_hiPS48_1", "g2_hiPS48_1")], breaks = bk, color = c(color_neg, color_pos), display_numbers = F,  cluster_cols = F, border_color = NA)
dev.off()
pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_cluster_0_1.pdf", height = 7, width = 7)
pheatmap(cluster_0_1_deg_df_dcast[upreg_both, c("g1_hiPS6_0", "g2_hiPS6_0", "g1_hiPS48_0", "g1_hiPS6_1", "g2_hiPS6_1", "g1_hiPS48_1", "g2_hiPS48_1")], breaks = bk, color = c(color_neg, color_pos), display_numbers = F,  cluster_cols = F, border_color = NA)
dev.off()
# # pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_cluster_0_only.pdf", height = 4, width = 7)
# pheatmap(cluster_0_1_deg_df_dcast[upreg_0_only, c("g1_hiPS6_0", "g2_hiPS6_0", "g1_hiPS48_0", "g1_hiPS6_1", "g2_hiPS6_1", "g1_hiPS48_1", "g2_hiPS48_1")], breaks = bk, color = c(color_neg, color_pos), display_numbers = F, cluster_rows = F, cluster_cols = F, border_color = NA)
# # dev.off()
# # pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_cluster_1_only.pdf", height = 7, width = 7)
# pheatmap(cluster_0_1_deg_df_dcast[upreg_1_only, c("g1_hiPS6_0", "g2_hiPS6_0", "g1_hiPS48_0", "g1_hiPS6_1", "g2_hiPS6_1", "g1_hiPS48_1", "g2_hiPS48_1")], breaks = bk, color = c(color_neg, color_pos), display_numbers = F, cluster_rows = F, cluster_cols = F, scale = "row", border_color = NA)
# # dev.off()
```

## Commonly down / up regulated genes: Overrepresentation test

### Examples of upregulated genes among all experiments which are core enrichment in these terms

```{r}
universe <- rownames(crispri_org)
set.seed(7)
library(clusterProfiler)
library(org.Hs.eg.db)

go_enrich_upreg_cluster_0_1 <- enrichGO(gene = upreg_both, 
                                        OrgDb = org.Hs.eg.db, 
                                        universe = universe,
                                        keyType = 'SYMBOL', 
                                        readable = T,
                                        ont = "ALL",
                                        pvalueCutoff = 0.05, 
                                        qvalueCutoff = 0.10)

go_enrich_downreg_cluster_0_1 <- enrichGO(gene = downreg_both, 
                                        OrgDb = org.Hs.eg.db, 
                                        universe = universe,
                                        keyType = 'SYMBOL', 
                                        readable = T,
                                        ont = "ALL",
                                        pvalueCutoff = 0.05, 
                                        qvalueCutoff = 0.10)

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/goenrich_upreg_cluster_0_1_cluster.pdf", height = 15, width = 7)
dotplot(go_enrich_upreg_cluster_0_1, showCategory = 50) + ggtitle("Upregulated genes in cluster 0 and 1")
# dev.off()
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/goenrich_downreg_cluster_0_1_cluster.pdf", height = 15, width = 7)
dotplot(go_enrich_downreg_cluster_0_1, showCategory = 50) + ggtitle("Downregulated genes in cluster 0 and 1")
# dev.off()

# saveRDS(list("upreg_cluster_0_1" = go_enrich_upreg_cluster_0_1@result,
#                 "downreg_cluster_0_1" = go_enrich_downreg_cluster_0_1@result),
#         "/Volumes/MyPassport/CRISPRi_L1s/10X/src/r_scripts/goenrich_cluster_0_1_cluster.rds")

go_cluster_0_1 <- readRDS("/Volumes/MyPassport/CRISPRi_L1s/10X/src/r_scripts/goenrich_cluster_0_1_cluster.rds")
# 
# write.xlsx(list("upreg_cluster_0_1" = go_enrich_upreg_cluster_0_1@result,
#                 "downreg_cluster_0_1" = go_enrich_downreg_cluster_0_1@result), "/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/goenrich_cluster_0_1_cluster.xlsx")
```


## Examples of upregulated genes among all experiments (violins)
```{r}
genes_enrichment_eg <- c("AKAP6", "NTRK2", "TOX", "SCN3A", "EPHA3", "GRIK1", "ADCY2", "FAT4", "UQCRFS1")#unique(unlist(str_split(go_enrich_upreg_cluster_0_1@result$geneID, "/")))
examples_crispri_upreg <- reshape2::melt(FetchData(crispri_org_join, c(genes_enrichment_eg,"cell_line", "condition", "seurat_clusters", "condition_cell_line_cluster_batch"))) %>% 
  # filter(value > 0) %>% 
  filter(seurat_clusters %in% c("0", "1"))
examples_crispri_upreg$condition <- factor(examples_crispri_upreg$condition, levels=c("LacZ", "gRNA1", "gRNA2"))
examples_crispri_upreg$seurat_clusters <- factor(examples_crispri_upreg$seurat_clusters, levels=c("0", "1"))

violin_padj <- function(df, gene){
  return(df %>% 
  filter(variable == gene) %>% 
  ggplot(aes(x=condition, y=value, fill=condition, color=condition)) + geom_jitter(height = 0, width = 0.2, alpha=0.7, size = 0.1) +
    geom_violin(alpha=0.5, color="black") + 
  # geom_boxplot(color="black", width=0.2, outliers = F) +
  facet_wrap(seurat_clusters~cell_line, ncol=9, scales = "free_x") +
   theme_bw()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    scale_color_manual(values = c("LacZ" = "grey",
                                  "gRNA1" = "#c8a4cd",
                                  "gRNA2" = "#9b67a3")) +
    scale_fill_manual(values = c("LacZ" = "grey",
                                  "gRNA1" = "#c8a4cd",
                                  "gRNA2" = "#9b67a3")) +
    ggtitle(gene) + labs(x="", y="Normalized Expression"))
}

deas_clusters$g1_hiPS6$`0`[genes_enrichment_eg,]
deas_clusters$g1_hiPS6$`1`[genes_enrichment_eg,]

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/violin_upreg_genes_cluster_0_1.pdf", height = 3)
violin_padj(examples_crispri_upreg, "AKAP6")
violin_padj(examples_crispri_upreg, "NTRK2")
violin_padj(examples_crispri_upreg, "TOX")
violin_padj(examples_crispri_upreg, "SCN3A")
violin_padj(examples_crispri_upreg, "EPHA3")
violin_padj(examples_crispri_upreg, "GRIK1")
violin_padj(examples_crispri_upreg, "ADCY2")
violin_padj(examples_crispri_upreg, "FAT4")
violin_padj(examples_crispri_upreg, "UQCRFS1")
# dev.off()

tmp <- subset(crispri_org_join, seurat_clusters %in% c("0", "1") & crispri_org_join$condition_cell_line %in% c("LacZ_hiPS48", "gRNA1_hiPS48"))

pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/vln_upreg_genes_cluster_0_1_examples_hiPS48_g1.pdf", height = 10, width = 4)
ggarrange(VlnPlot(tmp, "ADCY2", group.by = "seurat_clusters", split.by = "condition"),
VlnPlot(tmp, "TOX", group.by = "seurat_clusters", split.by = "condition"),
VlnPlot(tmp, "AKAP6", group.by = "seurat_clusters", split.by = "condition"), ncol=1, nrow=3)
dev.off()


```

## Number of nuclei per condition and cluster

```{r}
crispri_cluster_n <- reshape2::melt(FetchData(crispri_org_join, c("cell_line", "condition", "seurat_clusters", "condition_cell_line_cluster_batch")))

tmp <- crispri_cluster_n[,c("cell_line", "seurat_clusters", "condition")]
write.xlsx(as.data.frame(table(tmp$cell_line, tmp$seurat_clusters, tmp$condition)), "/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/crispri_org_nuclei_cluster_condition.xlsx")

```

## Percentage of nuclei expressing pluripotency markers
```{r}
clusters_per_cell <- FetchData(crispri_org, c("seurat_clusters", "condition"))
clusters_per_cell$barcode <- rownames(clusters_per_cell)
celltypes_per_cell <- merge(coldata_celltypes, clusters_per_cell, by.x="value", by.y="seurat_clusters")
colnames(celltypes_per_cell)[2] <- "celltype"
rownames(celltypes_per_cell) <- celltypes_per_cell$barcode
celltypes_per_cell$cluster_celltype_condition <- paste(celltypes_per_cell$value, paste(celltypes_per_cell$celltype, celltypes_per_cell$condition, sep="_"), sep="_")
crispri_org <- AddMetaData(crispri_org, metadata = celltypes_per_cell[,c("cluster_celltype_condition"), drop=F], col.name = "cluster_celltype_condition")

plt <- DotPlot(crispri_org, c("POU5F1", "SOX2", "NANOG" ), group.by = "cluster_celltype_condition")
pluripotent_markers <- plt$data[which(plt$data$features.plot != "SOX2"),c("id", "features.plot", "avg.exp", "pct.exp")]

pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/bar_perc_nuclei_pluripotency.pdf", height = 4, width = 6)
pluripotent_markers %>% 
  filter(features.plot == "NANOG") %>% 
  ggplot(aes(x=id, y=avg.exp*100, fill = avg.exp)) + geom_bar(stat="identity") + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(y="Percentage of nuclei in cluster\nexpressing NANOG (%)", x="", fill="Average\nExpression") + scale_y_continuous(limits=c(0,100), breaks = c(0,10,50,100))

pluripotent_markers %>% 
  filter(features.plot == "POU5F1") %>% 
  ggplot(aes(x=id, y=avg.exp*100, fill = avg.exp)) + geom_bar(stat="identity") + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(y="Percentage of nuclei in cluster\nexpressing POU5F1 (%)", x="", fill="Average\nExpression") + scale_y_continuous(limits=c(0,100), breaks = c(0,10,50,100))
dev.off()
```

