---
title: "L1-CRISPRi organoids: Cell typing"
output: html_notebook
---

This markdown relates to the visualization of day 15 cerebral organoids.

Main questions:
1. Cell type composition
2. Nuclei composition between conditions
3. Do L1-CRISPRi and control organoids significantly differ in cell type composition?
4. What are the consistently downregulated genes (mostly on-target)? Are they enriched in any specific biological processes?
5. What are the consistently upregulated genes (off-target)? Are they enriched in any specific biological processes?

## Read data and add metadata

```{r}
library(data.table)
library(Seurat)
library(tidyverse)
library(stringr)
library(openxlsx)
library(UpSetR)
library(RColorBrewer)
library(ggplot2)
library(ggpubr)

crispri_org <- readRDS("/Volumes/MyPassport/CRISPRi_L1s/10X/crispri.rds")

samples_metadata <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/metadata.tab", data.table = F)
metadata <- FetchData(crispri_org, "orig.ident")
metadata$barcode <- rownames(metadata)
metadata <- merge(metadata, samples_metadata, by.y="sample", by.x="orig.ident")
rownames(metadata) <- metadata$barcode
crispri_org <- AddMetaData(crispri_org, metadata[,c("guide", "batch", "cell_line", "seqnum")], col.name = c("condition", "batch", "cell_line", "seqnum"))

# Number of counts per nuclei
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/nCount_sample.pdf", height = 12)
VlnPlot(crispri_org, c("nCount_RNA"), group.by = "orig.ident") + theme(legend.position = "None")
# dev.off()

getPalette = colorRampPalette(brewer.pal(8, "Pastel1"))
crispri_org_split <- list()
dim_plots <- list()
for(i in unique(crispri_org$orig.ident)) crispri_org_split[[i]] <- subset(crispri_org, subset = orig.ident == i)
for(i in unique(crispri_org$orig.ident)) dim_plots[[i]] <- DimPlot(crispri_org_split[[i]], label = T, cols = getPalette(22)) & ggtitle(paste(unique(crispri_org_split[[i]]$cell_line), unique(crispri_org_split[[i]]$condition), "batch", unique(crispri_org_split[[i]]$batch))) & theme_bw() & theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank())

# UMAP per condition / batch
# 1. All nuclei together
# 2. Each experiment
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_org_L1CRISPRi.pdf", height = 4, width = 5)
DimPlot(crispri_org, label = T, cols = getPalette(22)) & ggtitle("All organoids samples") & theme_bw() & theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank())

dim_plots
# dev.off()

# Nuclei coming from hiPS6 CRISPRi using gRNA1
tmp <- subset(crispri_org, cell_line == "hiPS6" & condition %in% c("gRNA3", "LacZ"))
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_org_L1CRISPRi_g3.pdf", height = 4, width = 5)
DimPlot(tmp, label = T, cols = getPalette(22)) & ggtitle("All organoids samples") & theme_bw() & theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank())
# dev.off()

# Nuclei coming from hiPS48 CRISPRi using gRNA1
tmp <- subset(crispri_org, cell_line == "hiPS48" & condition %in% c("gRNA3", "LacZ"))
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_org_L1CRISPRi_hiPS48_g3_lacz.pdf", height = 4, width = 5)
DimPlot(tmp, label = T, cols = getPalette(22)) & ggtitle("All organoids samples (hiPS48)") & theme_bw() & theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank())
# dev.off()

# Nuclei coming from hiPS48 LacZ only
tmp <- subset(crispri_org, cell_line == "hiPS48" & condition %in% c("LacZ"))
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_org_L1CRISPRi_hiPS48_lacz.pdf", height = 4, width = 5)
DimPlot(tmp, label = T, cols = getPalette(22)) & ggtitle("LacZ hiPS48") & theme_bw() & theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank())
# dev.off()

# Nuclei coming from hiPS6 LacZ only
tmp <- subset(crispri_org, cell_line == "hiPS6" & condition %in% c("LacZ"))
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_org_L1CRISPRi_hiPS6_lacz.pdf", height = 4, width = 5)
DimPlot(tmp, label = T, cols = getPalette(22)) & ggtitle("LacZ hiPS6") & theme_bw() & theme(panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank())
# dev.off()
```

## Bio markers

Sort nuclei based on expression
```{r}
neuronal <- c("DCX")
deeplayer <- c("TBR1", "BCL11B")
interneurons <- c("GAD1", "DLX2")
cajalretzius <- c("RELN")
proliferating <- c("PAX6", "SOX2", "PCNA")
basalprog <- c("EOMES") 
microglia <- c("AIF1", "TMEM119")
markers <- unique(c(neuronal, deeplayer, interneurons, cajalretzius, proliferating, basalprog, microglia))

library(dplyr)
# Sorts nuclei based on expression so that we dont miss to see the expression by overlapping points
featureplot_sort_points <- function(df, gene, split = NA){
  df <- df %>% 
    drop_na(gene)
  minimum = min(df[,gene])
  maximum = max(df[,gene])
  
  plt <- df %>%
    arrange(get(gene)) %>%
    ggplot(aes(x=umap_1, y=umap_2, color=get(gene))) + geom_point(size=0.3) + theme_bw() + 
    scale_colour_gradient2(limits = c(minimum, maximum), high = "#1e386c", low = "#bfdbf0", mid = "#0394cb", midpoint = maximum/2) +
    labs(color = gene) +
    ggtitle(gene) +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold"))
  
  if(!is.na(split)){
    plt <- plt + facet_wrap(.~get(split), ncol=5)
  }
  return(plt)
}

markers_crispri_org <- FetchData(crispri_org, c("umap_1", "umap_2", "SOX2", "NES", "RBFOX3", "NCAM1", "PAX6", "EOMES", "FOXG1", "EMX1"))

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/crispri_org_markers_featureplot.pdf", width = 14, height = 6)
ggarrange(featureplot_sort_points(markers_crispri_org, "SOX2"),
featureplot_sort_points(markers_crispri_org, "NES"),
featureplot_sort_points(markers_crispri_org, "RBFOX3"),
featureplot_sort_points(markers_crispri_org, "NCAM1"),
featureplot_sort_points(markers_crispri_org, "PAX6"),
featureplot_sort_points(markers_crispri_org, "EOMES"),
featureplot_sort_points(markers_crispri_org, "FOXG1"),
featureplot_sort_points(markers_crispri_org, "EMX1"), ncol=4, nrow=2)
# dev.off()

# Using only hiPS6 CRISPRi gRNA1
tmp <- subset(crispri_org, cell_line == "hiPS6" & condition %in% c("gRNA3", "LacZ"))
examples_crispri_org <- FetchData(tmp, c("umap_1", "umap_2", "condition", "ENPEP", "CTNNA3", "KIAA1324L", "LINC00648", "NTRK2"))
```

## L1-derived gene examples (featureplots)
```{r}
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/crispri_org_examples_featureplot.pdf", width = 7, height = 3.5)
featureplot_sort_points(examples_crispri_org, "NTRK2", split = "condition")
featureplot_sort_points(examples_crispri_org, "KIAA1324L", split = "condition")
featureplot_sort_points(examples_crispri_org, "LINC00648", split = "condition")
featureplot_sort_points(examples_crispri_org, "ENPEP", split = "condition")
featureplot_sort_points(examples_crispri_org, "CTNNA3", split = "condition")
# dev.off()
```

## Examples of upregulated genes among all experiments (violins)
```{r}
examples_crispri_upreg <- reshape2::melt(FetchData(crispri_org_join, c("NTRK2", "PTX3", "PDE4B", "NR2E1", "SCN3A","cell_line", "condition", "seurat_clusters", "condition_cell_line_cluster_batch"))) %>% 
  filter(value > 0) %>% 
  filter(seurat_clusters %in% c("0", "1"))
examples_crispri_upreg$condition <- factor(examples_crispri_upreg$condition, levels=c("LacZ", "gRNA3", "gRNA1"))
examples_crispri_upreg$seurat_clusters <- factor(examples_crispri_upreg$seurat_clusters, levels=c("0", "1"))

violin_padj <- function(df, gene){
  return(df %>% 
  filter(variable == gene) %>% 
  ggplot(aes(x=condition, y=value, fill=condition, color=condition)) + geom_jitter(height = 0, width = 0.2, alpha=0.7, size = 0.1) +
    geom_violin(alpha=0.5, color="black") + 
  geom_boxplot(color="black", width=0.2, outliers = F) +
  facet_wrap(seurat_clusters~cell_line, ncol=9, scales = "free_x") +
   theme_bw()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    scale_color_manual(values = c("LacZ" = "grey",
                                  "gRNA3" = "#c8a4cd",
                                  "gRNA1" = "#9b67a3")) +
    scale_fill_manual(values = c("LacZ" = "grey",
                                  "gRNA3" = "#c8a4cd",
                                  "gRNA1" = "#9b67a3")) +
    ggtitle(gene) + labs(x="", y="Normalized Expression"))
}

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/violin_upreg_genes_cluster_0_1.pdf", height = 3)
violin_padj(examples_crispri_upreg, "NTRK2")
violin_padj(examples_crispri_upreg, "PTX3")
violin_padj(examples_crispri_upreg, "NR2E1")
violin_padj(examples_crispri_upreg, "SCN3A")
violin_padj(examples_crispri_upreg, "PDE4B")
# dev.off()
```

## Add metadata 
### Cell type
### Batch
### Condition
### Cell line
```{r}
coldata_celltypes <- reshape2::melt(list("NPCs" = as.character(c(0:2, 7)),
                    "Early neurons" = as.character(c(4,5)),
                    "Endothelial" = as.character(c(3,6))))

### ADD to metadata
clusters_per_cell <- FetchData(crispri_org, "seurat_clusters")
clusters_per_cell$barcode <- rownames(clusters_per_cell)
celltypes_per_cell <- merge(coldata_celltypes, clusters_per_cell, by.x="value", by.y="seurat_clusters")
colnames(celltypes_per_cell)[2] <- "celltype"
rownames(celltypes_per_cell) <- celltypes_per_cell$barcode
crispri_org <- AddMetaData(crispri_org, metadata = celltypes_per_cell[,c("celltype"), drop=F], col.name = "celltype")

cellline_per_cell <- FetchData(crispri_org, c("condition", "cell_line", "batch", "celltype"))
cellline_per_cell$barcode <- rownames(cellline_per_cell)
cellline_per_cell$condition_cell_line <- paste(cellline_per_cell$condition, cellline_per_cell$cell_line, sep="_")
cellline_per_cell$condition_celltype <- paste(cellline_per_cell$condition, cellline_per_cell$celltype, sep="_")
cellline_per_cell$condition_cell_line_batch <- paste(cellline_per_cell$condition, cellline_per_cell$cell_line, cellline_per_cell$batch, sep="_")
crispri_org <- AddMetaData(crispri_org, metadata = cellline_per_cell[,c("condition_cell_line", "condition_cell_line_batch", "condition_celltype"), drop=F], col.name = c("condition_cell_line", "condition_cell_line_batch", "condition_celltype"))
```

## Plot UMAP coloured by cell type
```{r}
# Only on hiPS6 LacZ (fig 5c)
tmp <- subset(crispri_org, cell_line == "hiPS6" & condition == "LacZ")
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_crispri_org_celltypes_lacz.pdf", width = 4.2, height = 3)
DimPlot(tmp, group.by = "celltype") & theme_bw() & 
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold")) & ggtitle("LacZ hiPS6 cell types")
# dev.off()

# Ratio of nuclei per cell type in hiPS6 LacZ
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/bar_ratio_crispri_org_celltypes_lacz.pdf", width = 4, height = 3)
FetchData(tmp, "celltype") %>% 
  ggplot(aes(x="LacZ", fill=celltype)) + geom_bar(stat="count", position="fill") + theme_bw() + labs(y="Ratio", x="", fill="Cell types")
# dev.off()

# Only on hiPS6 LacZ (supp fig 5b)
tmp <- subset(crispri_org, cell_line == "hiPS48" & condition == "LacZ")
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_crispri_org_celltypes_hiPS48_lacz.pdf", width = 4.2, height = 3)
DimPlot(tmp, group.by = "celltype") & theme_bw() & 
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold")) & ggtitle("LacZ hiPS48 cell types")
# dev.off()

# Ratio of nuclei per cell type in hiPS48 LacZ
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/bar_ratio_crispri_org_celltypes_hiPS48_lacz.pdf", width = 4, height = 3)
FetchData(tmp, "celltype") %>% 
  ggplot(aes(x="LacZ", fill=celltype)) + geom_bar(stat="count", position="fill") + theme_bw() + labs(y="Ratio", x="", fill="Cell types")
# dev.off()

```

## UMAP split or coloured by condition 
```{r}
# On hiPS6 experiments
tmp <- subset(crispri_org, cell_line == "hiPS6" & condition %in% c("LacZ", "gRNA3"))
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_crispri_org_celltypes_hips6.pdf", width = 6.7, height = 3)
DimPlot(tmp, group.by = "celltype", split.by = "condition") & theme_bw() & 
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold")) & ggtitle("hiPS6 cell types")
# dev.off()

# Coloured by condition
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_crispri_org_condition_cellcycle_hips6.pdf", width = 7, height = 6)
DimPlot(tmp, group.by = "condition", size = 0.1, cols = c("#794493", "grey"), alpha = 0.7) & theme_bw() & 
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold")) & ggtitle("hiPS6 conditions")
# dev.off()
```

## Cell type composition in each experiment
```{r}
nuclei_ratio <- FetchData(crispri_org, vars = c("celltype", "seurat_clusters","condition", "batch", "cell_line", "orig.ident", "condition_cell_line_batch"))
nuclei_ratio_g3 <- nuclei_ratio[which(nuclei_ratio$condition %in% c("gRNA3", "LacZ")),]

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/celltype_cluster_composition_bars.pdf", height = 4)
ggplot(nuclei_ratio_g3, aes(x=condition, fill=celltype)) + geom_bar(stat="count", position = "fill") + facet_wrap(cell_line~batch, scales = "free_x", ncol=6) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(x="Condition", y="Ratio of nuclei", fill = "Cell type")  + ggtitle("Cell type composition (split batches)")

ggplot(nuclei_ratio_g3, aes(x=condition, fill=celltype)) + geom_bar(stat="count", position = "fill") + facet_wrap(.~cell_line, scales = "free_x", ncol=6) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(x="Condition", y="Ratio of nuclei", fill = "Cell type")  + ggtitle("Cell type composition (pooled batches)")

ggplot(nuclei_ratio_g3, aes(x=condition, fill=seurat_clusters)) + geom_bar(stat="count", position = "fill") + facet_wrap(cell_line~batch, scales = "free_x", ncol=6) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(x="Condition", y="Ratio of nuclei", fill = "Cell type") + scale_fill_manual(values = brewer.pal(n=8,"Set2"))  + ggtitle("Cluster composition (split batches)")

ggplot(nuclei_ratio_g3, aes(x=condition, fill=seurat_clusters)) + geom_bar(stat="count", position = "fill") + facet_wrap(cell_line~batch, scales = "free_x", ncol=6) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(x="Condition", y="Ratio of nuclei", fill = "Cell type") + scale_fill_manual(values = brewer.pal(n=8,"Set2"))  + ggtitle("Cluster composition (split batches)")

ggplot(nuclei_ratio_g3, aes(x=condition, fill=seurat_clusters)) + geom_bar(stat="count", position = "fill") + facet_wrap(~cell_line, scales = "free_x", ncol=6) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(x="Condition", y="Ratio of nuclei", fill = "Cluster") + scale_fill_manual(values = brewer.pal(n=8,"Set2")) + ggtitle("Cluster composition (pooled batches)")

ggplot(nuclei_ratio_g3, aes(x=condition, fill=seurat_clusters)) + geom_bar(stat="count", position = "dodge") + facet_wrap(cell_line~batch, scales = "free_x", ncol=6) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(x="Condition", y="Ratio of nuclei", fill = "Cluster") + scale_fill_manual(values = brewer.pal(n=8,"Set2"))  + ggtitle("Cluster composition (split batches)")
# dev.off()
```

## Differential expression analysis 

Differences between LacZ and g1 - per cluster
```{r}
experiments <- list("g3_hiPS6" = c("LacZ_hiPS6_1", "LacZ_hiPS6_2", "LacZ_hiPS6_3", "LacZ_hiPS6_4",
                                   "gRNA3_hiPS6_1","gRNA3_hiPS6_2","gRNA3_hiPS6_3", "gRNA3_hiPS6_4"),
                    "g1_hiPS6" = c("LacZ_hiPS6_1", "LacZ_hiPS6_4",
                                   "gRNA3_hiPS6_1","gRNA1_hiPS6_4"),
                    "g3_hiPS48" = c("LacZ_hiPS48_1","LacZ_hiPS48_2", "gRNA3_hiPS48_1", "gRNA3_hiPS48_2"),
                    "g1_hiPS48" = c("LacZ_hiPS48_2", "gRNA1_hiPS48_2")) 

deas_clusters <- sapply(names(experiments), function(x) NULL)
clusters <- unique(crispri_org$seurat_clusters)
for(experiment in names(experiments)){
  deas_clusters[[experiment]] <- list()
  subset_crispri_org <- subset(crispri_org, condition_cell_line_batch %in% experiments[[experiment]])
  subset_crispri_org <- JoinLayers(subset_crispri_org)
  ident1 <- unique(subset_crispri_org$condition)[which(unique(subset_crispri_org$condition) != "LacZ")]
  for(cluster in clusters){
    if(!(experiment == "g3_hiPS48" & cluster == "7") & 
       !(experiment == "g3_hiPS48" & cluster == "5") &
       !(experiment == "g1_hiPS48" & cluster == "7")){
      print(experiment)
      print(cluster)
      print(ident1)
      deas_clusters[[experiment]][[cluster]] <- FindMarkers(subset_crispri_org, group.by = "condition", ident.1 = ident1, subset.ident = cluster, logfc.threshold = 0, only.pos = FALSE)
      deas_clusters[[experiment]][[cluster]]$gene <- rownames(deas_clusters[[experiment]][[cluster]])
    }
  }
}

deas_clusters_bound <- list()
for(experiment in names(experiments)){
  for(cluster in clusters){
    deas_clusters_bound[[paste(experiment, cluster, sep="_")]] <- deas_clusters[[experiment]][[cluster]]
  }
}

# write.xlsx(deas_clusters_bound, "/Volumes/MyPassport/CRISPRi_L1s/results/tables/genes_DEA.xlsx")

# saveRDS(deas_clusters, "/Volumes/MyPassport/CRISPRi_L1s/10X/r_scripts/src/crispri_org_deas_clusters.rds")
deas_clusters <- readRDS("/Volumes/MyPassport/CRISPRi_L1s/10X/src/r_scripts/crispri_org_deas_clusters.rds")
```

## Tag genes that have "same" or "opposite" direction to the results on hiPS6 g1

```{r}
deas_log2FC <- list()
for(i in as.character(0:7)){
  deas_log2FC[[i]] <- list()
  colnames(deas_clusters[["g3_hiPS6"]][[i]])[2] <- paste(i, "_g3_hiPS6_avg_log2FC", sep="")
  for(experiment in c("g3_hiPS48", "g1_hiPS6", "g1_hiPS48")){
    if(!(experiment == "g3_hiPS48" & i == "7") & !(experiment == "g3_hiPS48" & i == "5") & !(experiment == "g1_hiPS48" & i == "7")){
      deas_cluster_experiment <- deas_clusters[[experiment]][[i]]
      colnames(deas_cluster_experiment)[2] <- paste(i, experiment, "avg_log2FC", sep="_")
      
      genes <- rownames(deas_clusters[["g3_hiPS6"]][[i]])[which(rownames(deas_clusters[["g3_hiPS6"]][[i]]) %in% rownames(deas_cluster_experiment))]
      deas_log2FC[[i]][[experiment]] <- cbind(deas_clusters[["g3_hiPS6"]][[i]][genes, paste(i, "_g3_hiPS6_avg_log2FC", sep=""), drop=F],
                                              deas_cluster_experiment[genes, paste(i, experiment, "avg_log2FC", sep="_"), drop=F])
      deas_log2FC[[i]][[experiment]]$gene <- rownames(deas_log2FC[[i]][[experiment]])
      deas_log2FC[[i]][[experiment]]$direction <- ifelse(deas_log2FC[[i]][[experiment]][,paste(i, "_g3_hiPS6_avg_log2FC", sep="")] < 0 & deas_log2FC[[i]][[experiment]][,paste(i, experiment, "avg_log2FC", sep="_")] < 0, 
                                                         ifelse(deas_log2FC[[i]][[experiment]][,paste(i, "_g3_hiPS6_avg_log2FC", sep="")] < -0.5 & deas_log2FC[[i]][[experiment]][,paste(i, experiment, "avg_log2FC", sep="_")] < -0.5, "down_0.5", "down"), 
                                                         ifelse(deas_log2FC[[i]][[experiment]][,paste(i, "_g3_hiPS6_avg_log2FC", sep="")] > 0 & deas_log2FC[[i]][[experiment]][,paste(i, experiment, "avg_log2FC", sep="_")] > 0, ifelse(deas_log2FC[[i]][[experiment]][,paste(i, "_g3_hiPS6_avg_log2FC", sep="")] > 0.5 & deas_log2FC[[i]][[experiment]][,paste(i, experiment, "avg_log2FC", sep="_")] > 0.5, "up_0.5", "up"), 
                                                         "opposite"))
    }
  }
}
```



