---
title: "L1-CRISPRi organoids: Cell cycle and commonly dysregulated genes"
output: html_notebook
---

This markdown relates to the visualization of day 15 cerebral organoids.

Main questions:
1. Cell cycle
2. Biomarkers 
3. Characterize the present clusters
4. Commonly down / up regulated genes overrepresentation test

## Read data and add metadata

Including previously defined cell types (see crispri_org.Rmd), batch, cell line, condition, etc.

```{r}
library(data.table)
library(openxlsx)
library(Seurat)
library(stringr)
library(ggpubr)
library(ggplot2)
library(ggpubr)

crispri_org <- readRDS("/Volumes/MyPassport/CRISPRi_L1s/10X/crispri.rds")

samples_metadata <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/metadata.tab", data.table = F)
metadata <- FetchData(crispri_org, "orig.ident")
metadata$barcode <- rownames(metadata)
metadata <- merge(metadata, samples_metadata, by.y="sample", by.x="orig.ident")
rownames(metadata) <- metadata$barcode
crispri_org <- AddMetaData(crispri_org, metadata[,c("guide", "batch", "cell_line", "seqnum")], col.name = c("condition", "batch", "cell_line", "seqnum"))
coldata_celltypes <- reshape2::melt(list("NPCs" = as.character(c(0:2, 7)),
                    "Early neurons" = as.character(c(4,5)),
                    "Endothelial" = as.character(c(3,6))))

crispri_org_join <- JoinLayers(crispri_org)

### ADD cell type to metadata
clusters_per_cell <- FetchData(crispri_org_join, "seurat_clusters")
clusters_per_cell$barcode <- rownames(clusters_per_cell)
celltypes_per_cell <- merge(coldata_celltypes, clusters_per_cell, by.x="value", by.y="seurat_clusters")
colnames(celltypes_per_cell)[2] <- "celltype"
rownames(celltypes_per_cell) <- celltypes_per_cell$barcode
crispri_org_join <- AddMetaData(crispri_org_join, metadata = celltypes_per_cell[,c("celltype"), drop=F], col.name = "celltype")

cellline_per_cell <- FetchData(crispri_org_join, c("condition", "cell_line", "batch", "celltype", "seurat_clusters"))
cellline_per_cell$barcode <- rownames(cellline_per_cell)
cellline_per_cell$condition_cell_line <- paste(cellline_per_cell$condition, cellline_per_cell$cell_line, sep="_")
cellline_per_cell$condition_celltype <- paste(cellline_per_cell$condition, cellline_per_cell$celltype, sep="_")
cellline_per_cell$condition_cell_line_batch <- paste(cellline_per_cell$condition, cellline_per_cell$cell_line, cellline_per_cell$batch, sep="_")
crispri_org_join <- AddMetaData(crispri_org_join, metadata = cellline_per_cell[,c("condition_cell_line", "condition_cell_line_batch", "condition_celltype"), drop=F], col.name = c("condition_cell_line", "condition_cell_line_batch", "condition_celltype"))
cellline_per_cell$condition_cell_line_cluster <- paste(cellline_per_cell$condition_cell_line, cellline_per_cell$seurat_clusters, sep = "_")
cellline_per_cell$condition_cell_line_batch_cluster <- paste(cellline_per_cell$condition_cell_line_batch, cellline_per_cell$seurat_clusters, sep = "_")
cellline_per_cell$condition_cell_line_celltype <- paste(cellline_per_cell$condition_cell_line, cellline_per_cell$celltype, sep = "_")
cellline_per_cell$condition_cell_line_batch_celltype <- paste(cellline_per_cell$condition_cell_line_batch, cellline_per_cell$celltype, sep = "_")
crispri_org_join <- AddMetaData(crispri_org_join, metadata = cellline_per_cell[,c("condition_cell_line_batch_cluster", "condition_cell_line_batch_celltype", "condition_cell_line_cluster", "condition_cell_line_celltype"), drop=F], col.name = c("condition_cell_line_batch_cluster", "condition_cell_line_batch_celltype", "condition_cell_line_cluster", "condition_cell_line_celltype"))
```

## Cell cycle

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
crispri_org_join <- CellCycleScoring(crispri_org_join, s.features = s.genes, g2m.features = g2m.genes, set.ident = F, assay="RNA")
crispri_org_join <- AddMetaData(crispri_org_join, metadata = ifelse(crispri_org_join$Phase == "G1", "non-cycling", "cycling"), col.name = "cellCycle")

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/dotplot_crispri_org_markers_clusters.pdf")
DotPlot(crispri_org_join, features = unique(c("G2M.Score","S.Score",  "TJP1",  "GLI3", "CUX2", "PAX6", "SOX2", "PCNA", "EOMES", "BCL11B", "MAP2", "NCAM1", "RBFOX3", "TBR1", "DCX", "GAD1")), group.by = "seurat_clusters", assay = "RNA") & theme(axis.text.x = element_text(angle=90, hjust=1)) & coord_flip()
# dev.off()

cellcycle_scores <- FetchData(tmp, c("umap_1", "umap_2","day", "G2M.Score", "S.Score", "condition", "seurat_clusters"))
cellcycle_scores <- merge(cellcycle_scores, FetchData(tmp, c("celltype", "condition_cell_line_batch")), by="row.names")
rownames(cellcycle_scores) <- cellcycle_scores$Row.names
cellcycle_scores <- cellcycle_scores[,-1]

cellcycle_scores$score <- cellcycle_scores$G2M.Score + cellcycle_scores$S.Score
cellcycle_scores <- cellcycle_scores[order(cellcycle_scores$seurat_clusters, cellcycle_scores$condition),]

# Featureplot coloured by cell cycle score
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_crispri_org_cellcycle_score_hiPS6.pdf", width = 5, height = 4)
ggplot(cellcycle_scores, aes(x=umap_1, y=umap_2, colour=score)) + geom_point(size=0.5)+
  scale_colour_gradient2(low="white", mid="lightgrey", high = "red") + theme_bw() +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold"))
# dev.off()


## Cell cycle visualization per experiment
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/umap_crispri_org_cellcycle_score.pdf", width = 14, height = 8)
ggarrange(
  ggplot(cellcycle_scores, aes(x=umap_1, y=umap_2, colour=G2M.Score)) + geom_point(size=0.5) + facet_wrap(.~condition_cell_line_batch, ncol = 8) + scale_colour_gradient2(low="white", mid="lightgrey", high = "red") + theme_bw() +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold")),
          
  ggplot(cellcycle_scores, aes(x=umap_1, y=umap_2, colour=S.Score)) + geom_point(size=0.5) + facet_wrap(.~condition_cell_line_batch, ncol = 8) + scale_colour_gradient2(low="white", mid="lightgrey", high = "red") + theme_bw() +
    theme(axis.line = element_line(colour = "lightgrey"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          plot.title = element_text(face = "bold")), ncol=1)
# dev.off()
```

## Markers per experiment per cell type

```{r}
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/dotplot_crispri_org_markers_celltypes_lacz.pdf", height = 6, width = 7)
DotPlot(subset(crispri_org_join, condition == "LacZ" & cell_line == "hiPS6"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org (all)")
DotPlot(subset(crispri_org_join, condition == "LacZ" & cell_line == "hiPS48"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS48 org (all)")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS6_1"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org batch 1")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS6_2"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org batch 2")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS6_3"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org batch 3")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS6_4"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS6 org batch 4")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS48_1"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS48 org batch 1")
DotPlot(subset(crispri_org_join, condition_cell_line_batch == "LacZ_hiPS48_2"), features = markers, group.by  = "celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("LacZ hiPS48 org batch 2")
# dev.off()

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/dotplot_crispri_org_markers_celltypes.pdf", height = 5, width = 7)
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_1", "gRNA3_hiPS6_1")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g3 batch 1")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_2", "gRNA3_hiPS6_2")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g3 batch 2")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_2", "gRNA1_hiPS6_2")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g1 batch 2")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_3", "gRNA3_hiPS6_3")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g3 batch 3")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_4", "gRNA3_hiPS6_4")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g3 batch 4")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS6_4", "gRNA1_hiPS6_4")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS6 org g1 batch 4")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS48_1", "gRNA3_hiPS48_1")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS48 org g3 batch 1")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS48_2", "gRNA3_hiPS48_2")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS48 org g3 batch 2")
DotPlot(subset(crispri_org_join, condition_cell_line_batch %in% c("LacZ_hiPS48_2", "gRNA1_hiPS48_2")), features = markers, group.by  = "condition_celltype", assay = "RNA", cols = "RdBu") & coord_flip() & ggtitle("hiPS48 org g1 batch 2")
# dev.off()
```

## Characterize the clusters (DEA per cluster)

Top 10 differentially expressed genes per cluster
```{r}
set.seed(10)
crispri_org_join_markers <- FindAllMarkers(crispri_org_join, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

gene_annotation <- fread("/Volumes/MyPassport/annotations/human/gencode/v38/gencode.v38.annotation.gene_names.tab", data.table=F, skip=1, header =F )[,-1]
colnames(gene_annotation) <- c("gene_name", "gene_type")
crispri_org_join_markers <- merge(crispri_org_join_markers, gene_annotation, by.x="gene", by.y="gene_name")
crispri_org_join_markers_top10 <- crispri_org_join_markers %>% 
  filter(gene_type == "protein_coding") %>% 
  filter(pct.1 > 0.4) %>% 
  group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC) %>% 
  arrange(cluster)

# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/crispri_org_markers_dotplot.pdf", height = 12, width = 7)
DotPlot(crispri_org, features = crispri_org_join_markers_top10$gene, group.by  = "seurat_clusters", assay = "RNA", cols = "RdBu") & coord_flip() 
# dev.off()

# write.xlsx(crispri_org_join_markers, "/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/crispri_org_markers_clusters.xlsx")

```

## Log2FC of consistently upregulated or downregulated genes

So, the files read here are based on the results of hiPS6 organoids. We need to find the genes that are:
* Only upregulated in cluster 0 and not in cluster 1 (<=0.1 or NA) in gRNA1, 2, and in hiPS48.
* Only upregulated in cluster 1 and not in cluster 0 (<=0.1 or NA) in gRNA1, 2, and in hiPS48.
* Only downregulated in cluster 0 and not in cluster 1 (>= -0.1 or NA) in gRNA1, 2, and in hiPS48.
* Only downregulated in cluster 1 and not in cluster 0 (>= -0.1 or NA) in gRNA1, 2, and in hiPS48.
* Both up or downregulated in cluster 0 and 1

Define if they are only upregulated in cluster 0
```{r}
deas_clusters <- readRDS("/Volumes/MyPassport/CRISPRi_L1s/10X/r_scripts/src/crispri_org_deas_clusters.rds")

upreg_0 <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/upreg_cluster_0.txt", header = F, data.table = F)
upreg_1 <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/upreg_cluster_1.txt", header = F, data.table = F)
downreg_0 <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/downreg_cluster_0.txt", header = F, data.table = F)
downreg_1 <- fread("/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/downreg_cluster_1.txt", header = F, data.table = F)

# Only upregulated in cluster 0 and not in cluster 1 (<=0.1 or NA) in gRNA1, 2, and in hiPS48.
upreg_0_only <- intersect(intersect(upreg_0[which(deas_clusters$g3_hiPS6$`0`[upreg_0$V1, "avg_log2FC"] > 0.5 &
                                                    (deas_clusters$g3_hiPS6$`1`[upreg_0$V1, "avg_log2FC"] <= 0.1 | is.na(deas_clusters$g3_hiPS6$`1`[upreg_0$V1, "avg_log2FC"]))), "V1"],
                                    upreg_0[which(deas_clusters$g1_hiPS6$`0`[upreg_0$V1, "avg_log2FC"] > 0.5 &
                                                    (deas_clusters$g1_hiPS6$`1`[upreg_0$V1, "avg_log2FC"] <= 0.1 | is.na(deas_clusters$g1_hiPS6$`1`[upreg_0$V1, "avg_log2FC"]))), "V1"]),
                          upreg_0[which(deas_clusters$g3_hiPS48$`0`[upreg_0$V1, "avg_log2FC"] > 0.5 &
                                          (deas_clusters$g3_hiPS48$`1`[upreg_0$V1, "avg_log2FC"] <= 0.1 | is.na(deas_clusters$g3_hiPS48$`1`[upreg_0$V1, "avg_log2FC"]))), "V1"])
# GADD45G, PCDHA3

# Only upregulated in cluster 1 and not in cluster 0 (<=0.1 or NA) in gRNA1, 2, and in hiPS48.
upreg_1_only <- intersect(intersect(upreg_1[which(deas_clusters$g3_hiPS6$`1`[upreg_1$V1, "avg_log2FC"] > 0.5 &
                                                    (deas_clusters$g3_hiPS6$`0`[upreg_1$V1, "avg_log2FC"] <= 0.1 | is.na(deas_clusters$g3_hiPS6$`0`[upreg_1$V1, "avg_log2FC"]))), "V1"],
                                    upreg_1[which(deas_clusters$g1_hiPS6$`1`[upreg_1$V1, "avg_log2FC"] > 0.5 &
                                                    (deas_clusters$g1_hiPS6$`0`[upreg_1$V1, "avg_log2FC"] <= 0.1 | is.na(deas_clusters$g1_hiPS6$`0`[upreg_1$V1, "avg_log2FC"]))), "V1"]),
                          upreg_1[which(deas_clusters$g3_hiPS48$`1`[upreg_1$V1, "avg_log2FC"] > 0.5 &
                                          (deas_clusters$g3_hiPS48$`0`[upreg_1$V1, "avg_log2FC"] <= 0.1 | is.na(deas_clusters$g3_hiPS48$`0`[upreg_1$V1, "avg_log2FC"]))), "V1"])
# RRH, LINC01994

# Only downregulated in cluster 0 and not in cluster 1 (>= -0.1 or NA) in gRNA1, 2, and in hiPS48.
downreg_0_only <- intersect(intersect(downreg_0[which(deas_clusters$g3_hiPS6$`0`[downreg_0$V1, "avg_log2FC"] < -0.5 &
                                                    (deas_clusters$g3_hiPS6$`1`[downreg_0$V1, "avg_log2FC"] >= -0.1 | is.na(deas_clusters$g3_hiPS6$`1`[downreg_0$V1, "avg_log2FC"]))), "V1"],
                                    downreg_0[which(deas_clusters$g1_hiPS6$`0`[downreg_0$V1, "avg_log2FC"] < -0.5 &
                                                    (deas_clusters$g1_hiPS6$`1`[downreg_0$V1, "avg_log2FC"] >= -0.1 | is.na(deas_clusters$g1_hiPS6$`1`[downreg_0$V1, "avg_log2FC"]))), "V1"]),
                          downreg_0[which(deas_clusters$g3_hiPS48$`0`[downreg_0$V1, "avg_log2FC"] < -0.5 &
                                          (deas_clusters$g3_hiPS48$`1`[downreg_0$V1, "avg_log2FC"] >= -0.1 | is.na(deas_clusters$g3_hiPS48$`1`[downreg_0$V1, "avg_log2FC"]))), "V1"])
# CCDC71L

# Only downregulated in cluster 1 and not in cluster 0 (>= -0.1 or NA) in gRNA1, 2, and in hiPS48.
downreg_1_only <- intersect(intersect(downreg_1[which(deas_clusters$g3_hiPS6$`1`[downreg_1$V1, "avg_log2FC"] < -0.5 &
                                                    (deas_clusters$g3_hiPS6$`0`[downreg_1$V1, "avg_log2FC"] >= -0.1 | is.na(deas_clusters$g3_hiPS6$`0`[downreg_1$V1, "avg_log2FC"]))), "V1"],
                                    downreg_1[which(deas_clusters$g1_hiPS6$`1`[downreg_1$V1, "avg_log2FC"] < -0.5 &
                                                    (deas_clusters$g1_hiPS6$`0`[downreg_1$V1, "avg_log2FC"] >= -0.1 | is.na(deas_clusters$g1_hiPS6$`0`[downreg_1$V1, "avg_log2FC"]))), "V1"]),
                          downreg_1[which(deas_clusters$g3_hiPS48$`1`[downreg_1$V1, "avg_log2FC"] < -0.5 &
                                          (deas_clusters$g3_hiPS48$`0`[downreg_1$V1, "avg_log2FC"] >= -0.1 | is.na(deas_clusters$g3_hiPS48$`0`[downreg_1$V1, "avg_log2FC"]))), "V1"])
# "AC020891.2" "AL022341.2" "AC105230.1" "ALOX5"      "AC005670.1" "TRHDE-AS1"  "SEPTIN4"    "SNX22"      "AL359851.1" "AC007608.3" "GCHFR"      "AC092445.1" "RALY-AS1"  "AC113404.1" "AC012146.1" "AL133335.1"

upreg_0$type <- ifelse(upreg_0$V1 %in% upreg_1$V1, "upreg_both", 
                       ifelse(upreg_0$V1 %in% upreg_0_only, "upreg_0", "upreg_both"))
upreg_1$type <- ifelse(upreg_1$V1 %in% upreg_0$V1, "upreg_both", 
                       ifelse(upreg_1$V1 %in% upreg_1_only, "upreg_1", "upreg_both"))
downreg_0$type <- ifelse(downreg_0$V1 %in% downreg_1$V1, "downreg_both", 
                         ifelse(downreg_0$V1 %in% downreg_0_only, "downreg_0", "downreg_both"))
downreg_1$type <- ifelse(downreg_1$V1 %in% downreg_0$V1, "downreg_both", 
                         ifelse(downreg_1$V1 %in% downreg_1_only, "downreg_1", "downreg_both"))

deg <- rbind(upreg_0,
             upreg_1,
             downreg_0,
             downreg_1)
deg <- unique(deg)
colnames(deg)[1] <- "deg"
rownames(deg) <- deg$deg

upreg_both <- deg[which(deg$type == "upreg_both"),"deg"]
downreg_both <- deg[which(deg$type == "downreg_both"),"deg"]

# Both up or downregulated in cluster 0 and 1
upreg_both <- upreg_both[which(deas_clusters$g3_hiPS6$`0`[upreg_both, "avg_log2FC"] > 0.5 &
                                  deas_clusters$g3_hiPS6$`1`[upreg_both, "avg_log2FC"] > 0.5 &
                                  deas_clusters$g1_hiPS6$`0`[upreg_both, "avg_log2FC"] > 0.5 &
                                  deas_clusters$g1_hiPS6$`1`[upreg_both, "avg_log2FC"] > 0.5 &
                                  deas_clusters$g3_hiPS48$`0`[upreg_both, "avg_log2FC"] > 0.5 &
                                  deas_clusters$g3_hiPS48$`1`[upreg_both, "avg_log2FC"] > 0.5)]

downreg_both <- downreg_both[which(deas_clusters$g3_hiPS6$`0`[downreg_both, "avg_log2FC"] < -0.5 &
                                  deas_clusters$g3_hiPS6$`1`[downreg_both, "avg_log2FC"] < -0.5 &
                                  deas_clusters$g1_hiPS6$`0`[downreg_both, "avg_log2FC"] < -0.5 &
                                  deas_clusters$g1_hiPS6$`1`[downreg_both, "avg_log2FC"] < -0.5 &
                                  deas_clusters$g3_hiPS48$`0`[downreg_both, "avg_log2FC"] < -0.5 &
                                  deas_clusters$g3_hiPS48$`1`[downreg_both, "avg_log2FC"] < -0.5)]

deg <- deg[which(deg$type %in% c("upreg_0", "upreg_1", "downreg_0", "downreg_1") | deg$deg %in% upreg_both | deg$deg %in% downreg_both),]

```


```{r}
g3_hiPS6_0_deg_df <- deas_clusters$g3_hiPS6$`0`[which(deas_clusters$g3_hiPS6$`0`$gene %in% deg$deg),]
g3_hiPS6_1_deg_df <- deas_clusters$g3_hiPS6$`1`[which(deas_clusters$g3_hiPS6$`1`$gene %in% deg$deg),]
g1_hiPS6_0_deg_df <- deas_clusters$g1_hiPS6$`0`[which(deas_clusters$g1_hiPS6$`0`$gene %in% deg$deg),]
g1_hiPS6_1_deg_df <- deas_clusters$g1_hiPS6$`1`[which(deas_clusters$g1_hiPS6$`1`$gene %in% deg$deg),]
g3_hiPS48_0_deg_df <- deas_clusters$g3_hiPS48$`0`[which(deas_clusters$g3_hiPS48$`0`$gene %in% deg$deg),]
g3_hiPS48_1_deg_df <- deas_clusters$g3_hiPS48$`1`[which(deas_clusters$g3_hiPS48$`1`$gene %in% deg$deg),]

g3_hiPS6_0_deg_df$type <- "g3_hiPS6_0"
g3_hiPS6_1_deg_df$type <- "g3_hiPS6_1"
g1_hiPS6_0_deg_df$type <- "g1_hiPS6_0"
g1_hiPS6_1_deg_df$type <- "g1_hiPS6_1"
g3_hiPS48_0_deg_df$type <- "g3_hiPS48_0"
g3_hiPS48_1_deg_df$type <- "g3_hiPS48_1"

cluster_0_1_deg_df <- rbind(g3_hiPS6_0_deg_df,
                            g3_hiPS6_1_deg_df,
                            g1_hiPS6_0_deg_df,
                            g1_hiPS6_1_deg_df,
                            g3_hiPS48_0_deg_df,
                            g3_hiPS48_1_deg_df)

# upreg_cluster_0_1_deg_df[is.na(upreg_cluster_0_1_deg_df$avg_log2FC), "avg_log2FC"] <- 0
cluster_0_1_deg_df <- unique(cluster_0_1_deg_df)
cluster_0_1_deg_df_dcast <- reshape2::dcast(cluster_0_1_deg_df[,c("avg_log2FC", "type", "gene")], formula = gene~type, value.var = "avg_log2FC")

cluster_0_1_deg_df_dcast[is.na(cluster_0_1_deg_df_dcast)] <- 0
rownames(cluster_0_1_deg_df_dcast) <- cluster_0_1_deg_df_dcast$gene
cluster_0_1_deg_df_dcast <- cluster_0_1_deg_df_dcast[,-1]

bk_neg <- seq(-20,0, length.out=50)
bk_pos <- seq(0,3.5, length.out=50)
bk <- c(bk_neg, bk_pos)
bk <- unique(bk)
color_pos <- colorRampPalette(c("white", "red"))(50)
color_neg <- colorRampPalette(c("blue", "white"))(50)


# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/downreg_genes_cluster_0_1.pdf", height = 15, width = 7)
pheatmap(cluster_0_1_deg_df_dcast[deg[which(startsWith(deg$type, "downreg_both")),"deg"], c("g3_hiPS6_0", "g1_hiPS6_0", "g3_hiPS48_0", "g3_hiPS6_1", "g1_hiPS6_1", "g3_hiPS48_1")], breaks = bk, color = c(color_neg, color_pos), display_numbers = F, cluster_rows = F, annotation_row = deg[,"type", drop=F], cluster_cols = F, border_color = NA)
# dev.off()
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_cluster_0_1.pdf", height = 7, width = 7)
pheatmap(cluster_0_1_deg_df_dcast[deg[which(startsWith(deg$type, "upreg_both")),"deg"], c("g3_hiPS6_0", "g1_hiPS6_0", "g3_hiPS48_0", "g3_hiPS6_1", "g1_hiPS6_1", "g3_hiPS48_1")], breaks = bk, color = c(color_neg, color_pos), display_numbers = F, cluster_rows = F, annotation_row = deg[,"type", drop=F], cluster_cols = F, border_color = NA)
# dev.off()
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_cluster_0_only.pdf", height = 4, width = 7)
pheatmap(cluster_0_1_deg_df_dcast[deg[which(endsWith(deg$type, "0")),"deg"], c("g3_hiPS6_0", "g1_hiPS6_0", "g3_hiPS48_0", "g3_hiPS6_1", "g1_hiPS6_1", "g3_hiPS48_1")], breaks = bk, color = c(color_neg, color_pos), display_numbers = F, cluster_rows = F, annotation_row = deg[,"type", drop=F], cluster_cols = F, gaps_col = 3, gaps_row = 2, border_color = NA)
# dev.off()
# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/upreg_genes_cluster_1_only.pdf", height = 7, width = 7)
pheatmap(cluster_0_1_deg_df_dcast[deg[which(endsWith(deg$type, "1")),"deg"], c("g3_hiPS6_0", "g1_hiPS6_0", "g3_hiPS48_0", "g3_hiPS6_1", "g1_hiPS6_1", "g3_hiPS48_1")], breaks = bk, color = c(color_neg, color_pos), display_numbers = F, cluster_rows = F, annotation_row = deg[,"type", drop=F], cluster_cols = F, gaps_col = 3, gaps_row = 2, scale = "row", border_color = NA)
# dev.off()
```

## Commonly down / up regulated genes: Overrepresentation test

### Examples of upregulated genes among all experiments which are core enrichment in these terms

```{r}
universe <- rownames(crispri_org)
set.seed(7)
library(clusterProfiler)
library(org.Hs.eg.db)

go_enrich_upreg_cluster_1 <- enrichGO(gene = c(deg[which(startsWith(deg$type, "upreg_both") | startsWith(deg$type, "upreg_1")),"deg"]), 
                                        OrgDb = org.Hs.eg.db, 
                                        universe = universe,
                                        keyType = 'SYMBOL', 
                                        readable = T,
                                        ont = "ALL",
                                        pvalueCutoff = 0.05, 
                                        qvalueCutoff = 0.10)

go_enrich_upreg_cluster_0 <- enrichGO(gene = c(deg[which(startsWith(deg$type, "upreg_both") | startsWith(deg$type, "upreg_0")),"deg"]), 
                                        OrgDb = org.Hs.eg.db, 
                                        universe = universe,
                                        keyType = 'SYMBOL', 
                                        readable = T,
                                        ont = "ALL",
                                        pvalueCutoff = 0.05, 
                                        qvalueCutoff = 0.10)


# pdf("/Volumes/MyPassport/CRISPRi_L1s/10X/results/plots/goenrich_upreg_cluster_0_1_cluster.pdf", height = 15, width = 7)
dotplot(go_enrich_upreg_cluster_0, showCategory = 50) + ggtitle("Upregulated genes in cluster 0")
dotplot(go_enrich_upreg_cluster_1, showCategory = 50) + ggtitle("Upregulated genes in cluster 1")
# dev.off()

# write.xlsx(list("upreg_cluster_0" = go_enrich_upreg_cluster_0@result,
#                 "upreg_cluster_1" = go_enrich_upreg_cluster_1@result), "/Volumes/MyPassport/CRISPRi_L1s/10X/results/tables/goenrich_cluster_0_1_cluster.xlsx")

VlnPlot(crispri_org_join, group.by = "condition_cell_line_cluster", features = c("AKAP6", "ADCY2", "PDE4B"))
```


