---
title: "DNA methylation over L1s"
output: html_notebook
---

# >6kbp L1HS-L1PA4 promoter methylation

## Read data

Format to have type (hiPS/organoid), cell line, coordinates from region in question, and chromosomes (1:22, X, Y)
```{r}
library(ggplot2)
library(stringr)
library(data.table)
library(tidyverse)
library(ggpubr)

L1_promoter_methyl <- fread("/Volumes/MyPassport/CRISPRi_L1s/results/tables/FL_L1HS_PA4_promoter.MN10_WT_Rep1_ontDNA_iPSC_hiPS6_human_basecalls_5mC_5hmC_hg38.sorted.cohort.segmeth.mc10.mr1.segplot_data.csv", data.table = F)

colnames(L1_promoter_methyl)[1] <- "position"
L1_promoter_methyl$type <- ifelse(grepl("d15",L1_promoter_methyl$sample), "organoid", "hiPSC")
L1_promoter_methyl$cell_line <- ifelse(grepl("hiPS48", L1_promoter_methyl$sample), "hiPS48",
                                       ifelse(grepl("hiPS10", L1_promoter_methyl$sample), "hiPS10",
                                              ifelse(grepl("hiPS6", L1_promoter_methyl$sample), "hiPS6", "other?")))
L1_promoter_methyl$sample_name <- paste(L1_promoter_methyl$type, L1_promoter_methyl$cell_line, sep="_")
L1_promoter_methyl$coords <- paste(sapply(str_split(L1_promoter_methyl$position, ":"), `[[`, 1), sapply(str_split(L1_promoter_methyl$position, ":"), `[[`, 2), sep=":")
L1_promoter_methyl$chr <- sapply(str_split(L1_promoter_methyl$position, ":"), `[[`, 1)
L1_promoter_methyl <- L1_promoter_methyl[which(L1_promoter_methyl$chr %in% c(paste("chr", 1:22, sep=""), "chrX", "chrY")), ]
```

## Visualize
L1 promoter methylation violins per cell line in hiPSC state and day 15 cerebral organoids.
```{r}
# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/l1hs_l1pa2_l1pa3_l1pa4_promoter_methylation.pdf", height = 4)
ggplot(L1_promoter_methyl, aes(x=cell_line, color=cell_line, y=modbase)) + geom_jitter(width = 0.2, height = 0, size=0.1) + 
  geom_violin(alpha=0.5) +
  geom_boxplot(width =0.2, alpha=0.5, outliers=F) +
  theme_bw() + facet_wrap(.~type, scales = "free_x") + 
  labs(x="", y="% methylation", color="Cell line") +
  ggtitle(">6kbp L1HS-L1PA4 promoter methylation")
# dev.off()
```

## L1 promoter methylation violins in hiPS6. 
### Comparison between hiPSC and day 15 cerebral organoids.
```{r}
# Just the hiPS that we have organoids on
L1_promoter_methyl <- L1_promoter_methyl[which(L1_promoter_methyl$cell_line != "hiPS10"),]

# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/l1hs_l1pa2_l1pa3_l1pa4_split_promoter_methylation_hiPS6_vs_hiPS6_org.pdf", height = 5, width = 5)
L1_promoter_methyl %>% 
  filter(cell_line == "hiPS6") %>% 
ggplot(aes(x=sample_name, y=modbase, color=sample_name)) + geom_jitter(width = 0.2, height = 0, size=0.1) + 
  geom_violin(alpha=0.5) +
  geom_boxplot(width =0.2, alpha=0.5, outliers=F) +
  theme_bw() + 
  facet_wrap(.~group, ncol=2) +
  labs(x="", y="% methylation", color="Cell line") +
  ggtitle(">6kbp L1HS-L1PA4 promoter methylation") + stat_compare_means(comparisons = list(c("hiPSC_hiPS6", "organoid_hiPS6"))) + theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          plot.title = element_text(face = "bold")) + scale_color_manual(values = c("hiPSC_hiPS6" = "#7ea3d6", 
                                                                                    "organoid_hiPS6" = "#7cab72"))
# dev.off()
```

### Comparison between hiPSC and day 15 cerebral organoids: Zoom-in.
```{r}
# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/l1hs_l1pa2_l1pa3_l1pa4_split_promoter_methylation_hiPS6_vs_hiPS6_org_zoom.pdf", height = 5, width = 5)
L1_promoter_methyl %>% 
  filter(cell_line == "hiPS6") %>% 
ggplot(aes(x=sample_name, y=modbase, color=sample_name)) + geom_jitter(width = 0.2, height = 0, size=0.1) + 
  theme_bw() + 
  facet_wrap(.~group, ncol=2) +
  labs(x="", y="% methylation", color="Cell line") +
  ggtitle(">6kbp L1HS-L1PA4 promoter methylation") + stat_compare_means(comparisons = list(c("hiPSC_hiPS6", "organoid_hiPS6"))) + theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          plot.title = element_text(face = "bold")) + scale_color_manual(values = c("hiPSC_hiPS6" = "#7ea3d6", 
                                                                                    "organoid_hiPS6" = "#7cab72")) +
  stat_summary(fun.y=mean,col='red',geom='point') +
  coord_cartesian(ylim = c(0.65,0.95)) + # Here is the zoom in
    theme(text = element_text(size = 20), title = element_text(size=15)) 
# dev.off()
```

## DNA methylation status over L1 promoters marked with H3K4me3
```{r}
# Set order of violins
L1_promoter_methyl$cell_line <- factor(L1_promoter_methyl$cell_line, levels = c("hiPS6", "hiPS48"))
L1_h3k4me3_peak <- fread("/Volumes/MyPassport/CRISPRi_L1s/results/tables/H3K4me3_tagdir_regions_merged_hiPS6_hiPS48_intersect_FL_L1PA_promoters.bed", skip=1, header = F)
L1_h3k4me3_peak$coords <- paste(L1_h3k4me3_peak$V1, paste(L1_h3k4me3_peak$V2, L1_h3k4me3_peak$V3, sep="-"), sep=":")

L1_promoter_methyl$h3k4me3_peak <- ifelse(L1_promoter_methyl$coords %in% L1_h3k4me3_peak$coords, "H3K4me3+ ", "")
L1_promoter_methyl$h3k4me3_peak_cell_line <- paste(L1_promoter_methyl$h3k4me3_peak, L1_promoter_methyl$cell_line, sep="")
L1_promoter_methyl$h3k4me3_peak_cell_line <- factor(L1_promoter_methyl$h3k4me3_peak_cell_line, levels = c("hiPS6", "H3K4me3+ hiPS6", "hiPS48", "H3K4me3+ hiPS48"))
# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/l1hs_l1pa2_l1pa3_l1pa4_split_promoter_methylation_hiPS6_hiPS48.pdf", height = 4, width = 10)
ggplot(L1_promoter_methyl[which(L1_promoter_methyl$type == "hiPSC"),], aes(x=h3k4me3_peak_cell_line, color=h3k4me3_peak_cell_line, y=modbase)) + geom_jitter(width = 0.2, height = 0, size=0.1) + 
  geom_violin(alpha=0.5) +
  geom_boxplot(width =0.2, alpha=0.5, outliers=F) +
  theme_bw() + 
  facet_wrap(.~group, ncol=4) +
  labs(x="", y="% methylation", color="Cell line") +
  ggtitle(">6kbp L1HS-L1PA4 promoter methylation") + theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(face = "bold")) + scale_color_manual(values = c("hiPS6" = "#7ea3d6", 
                                                                                    "hiPS48" = "#86bbcc",
                                                                                    "H3K4me3+ hiPS6" = "#7ea3d6", 
                                                                                    "H3K4me3+ hiPS48"= "#86bbcc"))
# dev.off()
```

## DNA methylation correlation with L1 transcription

```{r}
# Load TE normalized counts
load("/Volumes/MyPassport/CRISPRi_L1s/bulk/src/r_scripts/TE_uniq_DEA.RData")

# Subset to only include >6kbp L1HS-L1PA4
FL_L1PA_bed <- te_counts[rownames(FL_L1PA_norm), c("Chr", "Start", "End", "Strand", "Geneid")]
FL_L1PA_bed <- FL_L1PA_bed[which(grepl("L1HS|L1PA2|L1PA3|L1PA4", FL_L1PA_bed$Geneid)), ]
FL_L1PA_bed$tmp_start <- ifelse(FL_L1PA_bed$Strand == "+", FL_L1PA_bed$Start, FL_L1PA_bed$End - 900)
FL_L1PA_bed$tmp_end <- ifelse(FL_L1PA_bed$Strand == "+", FL_L1PA_bed$Start + 900, FL_L1PA_bed$End)
FL_L1PA_bed$coords <- paste(FL_L1PA_bed$Chr, paste(FL_L1PA_bed$tmp_start, FL_L1PA_bed$tmp_end, sep = "-"), sep=":")
FL_L1PA_bed <- FL_L1PA_bed[which(FL_L1PA_bed$Chr %in% c(paste("chr", 1:22, sep=""), "chrX", "chrY")), ]
# These are the normalized counts
FL_L1PA_norm$TE_id <- rownames(FL_L1PA_norm)

library(ggpmisc)
#### hiPS6 ####
hiPS6_samples <- samplesheet[which(startsWith(samplesheet$samples, "DA") & grepl("h6", samplesheet$samples)),"samples"]
# Coordinates of elements which we have the methylation data on
FL_L1PA_coord_methyl <- FL_L1PA_bed[which(FL_L1PA_bed$coords %in% L1_promoter_methyl$coords),"Geneid"]
# Subset the normalized counts for hiPS6
FL_L1PA_norm_found_ont_hiPS6 <- FL_L1PA_norm[FL_L1PA_coord_methyl, c(hiPS6_samples, "TE_id")]
# Melt to reformat the data frame
FL_L1PA_norm_found_ont_hiPS6_melt <- reshape2::melt(FL_L1PA_norm_found_ont_hiPS6)
# Get the coordinates of these elements
FL_L1PA_norm_found_ont_hiPS6_melt <- merge(FL_L1PA_norm_found_ont_hiPS6_melt, FL_L1PA_bed[,c("Geneid", "coords")], by.x = "TE_id", by.y = "Geneid")
# Merge with the methylation data of hiPS6
L1_promoter_methyl_hiPS6 <- L1_promoter_methyl[which(L1_promoter_methyl$type == "hiPSC" & L1_promoter_methyl$cell_line == "hiPS6"), c("coords", "modbase")]
FL_L1PA_norm_found_ont_hiPS6_melt <- merge(FL_L1PA_norm_found_ont_hiPS6_melt, L1_promoter_methyl_hiPS6)

#### hiPS48 ####
hiPS48_samples <- c(samplesheet[which(startsWith(samplesheet$samples, "DA") & grepl("h48", samplesheet$samples)),"samples"], "TE_id")
FL_L1PA_norm_found_ont_hiPS48 <- FL_L1PA_norm[FL_L1PA_coord_methyl, hiPS48_samples]
FL_L1PA_norm_found_ont_hiPS48_melt <- reshape2::melt(FL_L1PA_norm_found_ont_hiPS48)
FL_L1PA_norm_found_ont_hiPS48_melt <- merge(FL_L1PA_norm_found_ont_hiPS48_melt, FL_L1PA_bed[,c("Geneid", "coords")], by.x = "TE_id", by.y = "Geneid")
L1_promoter_methyl_hiPS48 <- L1_promoter_methyl[which(L1_promoter_methyl$type == "hiPSC" & L1_promoter_methyl$cell_line == "hiPS48"), c("coords", "modbase")]
FL_L1PA_norm_found_ont_hiPS48_melt <- merge(FL_L1PA_norm_found_ont_hiPS48_melt, L1_promoter_methyl_hiPS48)

# Create a column with TE subfamily name
FL_L1PA_norm_found_ont_hiPS6_melt$TE_subfamily <- sapply(str_split(FL_L1PA_norm_found_ont_hiPS6_melt$TE_id, "_dup"), `[[`, 1)
FL_L1PA_norm_found_ont_hiPS48_melt$TE_subfamily <- sapply(str_split(FL_L1PA_norm_found_ont_hiPS48_melt$TE_id, "_dup"), `[[`, 1)

# Finally visualize
# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/corr_methylation_expression_hiPS.pdf")
ggplot(FL_L1PA_norm_found_ont_hiPS6_melt, aes(x=log2(value+1), y=modbase)) + geom_point(size=0.3, alpha = 0.5, color = "#7da2d5") + 
  theme_bw() + 
  # geom_smooth(method = "lm", formula= y~x) + 
  stat_poly_line(color = "black", size=0.5) +
  stat_poly_eq(use_label(c("adj.R2", "p"))) +
  facet_wrap(variable~TE_subfamily, ncol=4) + labs(x="log2(Normalized Expression + 1)", y="% Methylation") + ggtitle("hiPS6 >6kbp L1HS-PA4 methylation vs expression") +
  theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          plot.title = element_text(face = "bold"))

ggplot(FL_L1PA_norm_found_ont_hiPS48_melt, aes(x=log2(value+1), y=modbase)) + geom_point(size=0.3, alpha = 0.5, color = "#7da2d5") + 
  theme_bw() + 
  # geom_smooth(method = "lm", formula= y~x) + 
  stat_poly_line(color = "black", size=0.5) +
  stat_poly_eq(use_label(c("adj.R2", "p"))) +
  facet_wrap(variable~TE_subfamily, ncol=4) + labs(x="log2(Normalized Expression + 1)", y="% Methylation") + ggtitle("hiPS48 >6kbp L1HS-PA4 methylation vs expression") +
  theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "lightgrey", fill=NA),
          panel.background = element_blank(),
          plot.title = element_text(face = "bold"))
# dev.off()

```

## Elements only found to be expressed in hiPS
```{r}
L1_promoter_methyl_hips_only <- fread("/Volumes/MyPassport/CRISPRi_L1s/results/tables/FL_L1PA_WT_hiPSC_only_expressed_promoter.MN10_WT_Rep1_ontDNA_iPSC_hiPS6_human_basecalls_5mC_5hmC_hg38.sorted.cohort.segmeth.tsv", data.table = F)

colnames(L1_promoter_methyl_hips_only)[1] <- "position"
L1_promoter_methyl_hips_only <- L1_promoter_methyl_hips_only[,c(1:6, which(grepl("m_methfrac", colnames(L1_promoter_methyl_hips_only))))]

# Add information on cell line and type of sample (hiPS/org)
L1_promoter_methyl_hips_only_melt <- reshape2::melt(L1_promoter_methyl_hips_only[,c(7:ncol(L1_promoter_methyl_hips_only), 1)])
L1_promoter_methyl_hips_only_melt$type <- ifelse(grepl("hiPS48", L1_promoter_methyl_hips_only_melt$variable), "hiPS48",
                                                 ifelse(grepl("hiPS6", L1_promoter_methyl_hips_only_melt$variable), 
                                                        ifelse(grepl("org_hiPS6", L1_promoter_methyl_hips_only_melt$variable), "hiPS6 org", "hiPS6"), "other?"))

library(ggpubr)
# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/l1pa_promoter_methylation_expressed_hiPS_only.pdf", height = 6, width = 5)
ggplot(L1_promoter_methyl_hips_only_melt[which(L1_promoter_methyl_hips_only_melt$type != "hiPS48"),], aes(x=type, y=value, fill=type)) + geom_jitter(aes(color=type),alpha=0.2, height = 0, width = 0.2) + geom_violin(alpha=0.3) + geom_boxplot(width=0.2, outliers = F) + theme_bw() + labs(y="% methylation", x="", fill="", color="") + scale_fill_manual(values = c("#56a5b2", "#87c872")) + scale_color_manual(values= c("#56a5b2", "#87c872")) + stat_compare_means(comparisons = list(c("hiPS6", "hiPS6 org")), label.y = 1) + ggtitle(">6kbp L1PAs only\nexpressed in hiPS") + lims(y=c(0.2,1.05)) + theme(text = element_text(size = 20))
# dev.off()
```

## Methylation of elements found to be expressed in both hiPS and organoids

Here we compare the three groups (expressed only in hiPS, only in organoids, and in both)
```{r}
L1_promoter_methyl_org_hips <- fread("/Volumes/MyPassport/CRISPRi_L1s/results/tables/FL_L1PA_LacZ_org_WT_hiPSC_expressed_promoter.MN10_WT_Rep1_ontDNA_iPSC_hiPS6_human_basecalls_5mC_5hmC_hg38.sorted.cohort.segmeth.tsv", data.table = F)

colnames(L1_promoter_methyl_org_hips)[1] <- "position"
L1_promoter_methyl_org_hips <- L1_promoter_methyl_org_hips[,c(1:6, which(grepl("m_methfrac", colnames(L1_promoter_methyl_org_hips))))]

# Last time, add info
L1_promoter_methyl_org_hips_melt <- reshape2::melt(L1_promoter_methyl_org_hips[,c(7:ncol(L1_promoter_methyl_org_hips), 1)])
L1_promoter_methyl_org_hips_melt$type <- ifelse(grepl("hiPS48", L1_promoter_methyl_org_hips_melt$variable), "hiPS48",
                                                 ifelse(grepl("hiPS6", L1_promoter_methyl_org_hips_melt$variable), 
                                                        ifelse(grepl("org_hiPS6", L1_promoter_methyl_org_hips_melt$variable), "hiPS6 org", "hiPS6"), "other?"))

# Add column per data frame with the group of elements they belong to
L1_promoter_methyl_hips_only_melt$expressed <- "hiPS_only"
L1_promoter_methyl_org_hips_melt$expressed <- "hiPS_org"

# Bind them
L1_promoter_methyl_melt <- rbind(L1_promoter_methyl_hips_only_melt,
                                 L1_promoter_methyl_org_hips_melt)

L1_promoter_methyl_melt$type_expressed <- paste(L1_promoter_methyl_melt$type, L1_promoter_methyl_melt$expressed, sep="_")

# Compare all conditions
combinations <- combn(unique(L1_promoter_methyl_melt[which(L1_promoter_methyl_melt$type != "hiPS48"),"type_expressed"]), 2)
# Set the order right
L1_promoter_methyl_melt$expressed <- factor(L1_promoter_methyl_melt$expressed, levels = c("hiPS_org", "hiPS_only"))

# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/l1pa_promoter_methylation_expressed_venn.pdf", height = 5, width = 6)
L1_promoter_methyl_melt[which(L1_promoter_methyl_melt$type != "hiPS48"),] %>% # Only hiPS6
  drop_na() %>% 
ggplot(aes(x=expressed, y=value, fill=expressed)) + geom_point(aes(color=expressed), alpha=0.2, pch = 21, position = position_jitterdodge(dodge.width = 0.9, jitter.height = 0, jitter.width = 0.1)) + geom_violin(alpha=0.3) + geom_boxplot(width=0.2, outliers = F, position = position_dodge(width = 0.9)) + theme_bw() + 
  labs(y="% methylation", x="", fill="", color="") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + facet_wrap(.~type) +
  scale_fill_manual(values = c("#5bd641", "#32b2a5", "#2c8cdb")) +
  scale_color_manual(values= c("#aad3a2", "#87b3b0", "#b0c7dd")) +
  ggtitle(">6kbp L1PAs") + 
    theme(text = element_text(size = 20)) 
# dev.off()

```


## Zoom in to the differences between the groups
```{r}
# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/l1pa_promoter_methylation_expressed_venn_pvals_zoom.pdf", height = 3, width = 5)
L1_promoter_methyl_melt[which(L1_promoter_methyl_melt$type == "hiPS6 org"),] %>% # Only organoids
  drop_na() %>% 
ggplot(aes(x=expressed, y=value, fill=expressed)) + geom_point(aes(color=expressed), alpha=0.2, pch = 21, position = position_jitterdodge(dodge.width = 0.9, jitter.height = 0, jitter.width = 0.7)) + theme_bw() + 
  stat_summary(fun.y=mean,col='red',geom='point') +
   stat_summary(geom = "line", fun = mean, group = 1, linetype="dashed") +
labs(y="% methylation", x="", fill="", color="") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#5bd641", "#32b2a5", "#2c8cdb")) +
  scale_color_manual(values= c("#aad3a2", "#87b3b0", "#b0c7dd")) +
  ggtitle(">6kbp L1PAs: Organoids hiPS6") + 
  coord_cartesian(ylim = c(0.85,1)) + # Here is the zoom in
    theme(text = element_text(size = 20), title = element_text(size=15)) 


L1_promoter_methyl_melt[which(L1_promoter_methyl_melt$type == "hiPS6"),] %>% # Only in hiPS6
  drop_na() %>% 
ggplot(aes(x=expressed, y=value, fill=expressed)) + geom_point(aes(color=expressed), alpha=0.2, pch = 21, position = position_jitterdodge(dodge.width = 0.9, jitter.height = 0, jitter.width = 0.7)) + theme_bw() + 
  stat_summary(fun.y=mean,col='red',geom='point') +
   stat_summary(geom = "line", fun = mean, group = 1, linetype="dashed") +
labs(y="% methylation", x="", fill="", color="") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#5bd641", "#32b2a5", "#2c8cdb")) +
  scale_color_manual(values= c("#aad3a2", "#87b3b0", "#b0c7dd")) +
  ggtitle(">6kbp L1PAs: hiPS6") + 
  coord_cartesian(ylim = c(0.70,1)) + # Zoom in
    theme(text = element_text(size = 20), title = element_text(size=15)) 
# dev.off()
```


```{r}
# pdf("/Volumes/MyPassport/CRISPRi_L1s/results/plots/l1pa_promoter_methylation_expressed_venn_pvals_wilcox_hiPS_only_both.pdf", height = 5, width = 5)
L1_promoter_methyl_melt[which(L1_promoter_methyl_melt$type == "hiPS6 org"),] %>% # Organoid samples
  drop_na() %>% 
ggplot(aes(x=expressed, y=value, fill=expressed)) + geom_point(aes(color=expressed), alpha=0.2, pch = 21, position = position_jitterdodge(dodge.width = 0.9, jitter.height = 0, jitter.width = 0.1)) + geom_violin(alpha=0.3) + geom_boxplot(width=0.2, outliers = F, position = position_dodge(width = 0.9)) + theme_bw() + 
  labs(y="% methylation", x="", fill="", color="") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#32b2a5", "#2c8cdb")) +
  scale_color_manual(values= c("#87b3b0", "#b0c7dd")) +
  ggtitle(">6kbp L1PAs: Organoids hiPS6") + 
    theme(text = element_text(size = 20), title = element_text(size=15)) + stat_compare_means()

L1_promoter_methyl_melt[which(L1_promoter_methyl_melt$type == "hiPS6"),] %>% # hiPS samples
  drop_na() %>% 
ggplot(aes(x=expressed, y=value, fill=expressed)) + geom_point(aes(color=expressed), alpha=0.2, pch = 21, position = position_jitterdodge(dodge.width = 0.9, jitter.height = 0, jitter.width = 0.1)) + geom_violin(alpha=0.3) + geom_boxplot(width=0.2, outliers = F, position = position_dodge(width = 0.9)) + theme_bw() + 
  labs(y="% methylation", x="", fill="", color="") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("#32b2a5", "#2c8cdb")) +
  scale_color_manual(values= c("#87b3b0", "#b0c7dd")) +
  ggtitle(">6kbp L1PAs: hiPS6") + 
    theme(text = element_text(size = 20), title = element_text(size=15)) + stat_compare_means()
# dev.off()
```

